---
title: "Primary Headwater Stream Analyses"
author: "Nathan Byer"
date: '`r Sys.Date()`'
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    fig_width: 6
    fig_height: 4
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width = 600)

library(AICcmodavg)
library(broom)
library(camtrapR)
library(caret)
library(chron)
library(cooccur)
library(dismo)
library(dplyr)
library(ggmap)
library(ggplot2)
library(gstat)
library(igraph)
library(indicspecies)
library(lme4)
library(lubridate)
library(MASS)
library(raster)
library(RColorBrewer)
library(rgdal)
library(rjags)
library(shiny)
library(sp)
library(spOccupancy)
library(terra)
library(tibble)
library(tidyr)
library(tidyverse)
library(vegan)
library(visNetwork)
library(ggord)
```

# Reading in data

```{r, warning = FALSE, message = FALSE}
phwh<-read.csv("PHWH_postgresexport.csv",header=T,na.strings=c("","NA"))


```

# creating summary files for downstream analysis

```{r, warning = FALSE, message= FALSE}

sitechecks<-data.frame(phwh$stream_name,phwh$date)

colnames(sitechecks)<-c("stream_names","date")

sitechecks$date<- as.Date(sitechecks$date , format = "%Y/%m/%d")

sitechecks$datecount<-as.numeric(sitechecks$date-min(sitechecks$date)+1)

sitechecks_surveyhistory<-sitechecks %>%
  group_by(stream_names) %>% 
  mutate(check=as.character(dense_rank(date))) %>%
  arrange(stream_names) %>% 
  distinct(stream_names, datecount, .keep_all = TRUE) %>%
  spread(check,datecount) %>%
  summarise(`1` = mean(`1`,na.rm=T),
            `2`=mean(`2`,na.rm=T),
            `3`=mean(`3`,na.rm=T),
            `4`=mean(`4`,na.rm=T),
            `5`=mean(`5`,na.rm=T),
            `6`=mean(`6`,na.rm=T),
            `7`=mean(`7`,na.rm=T),
            `8`=mean(`8`,na.rm=T),
            `9` = mean(`9`,na.rm=T),
            `10`=mean(`10`,na.rm=T),
            `11`=mean(`11`,na.rm=T),
            `12`=mean(`12`,na.rm=T),
            `13`=mean(`13`,na.rm=T),
            `14`=mean(`14`,na.rm=T),
            `15`=mean(`15`,na.rm=T),
            `16`=mean(`16`,na.rm=T))

phwh_subset<-phwh %>%
  dplyr::select(stream_name,reservation,river_basin,latitude,longitude,date,substrate_metric_total,percent_best_types,per_boulder_slabs,per_boulder_256mm,per_bedrock,per_cobble,per_gravel,per_sand,per_silt,per_leaf_woody,per_detritus,per_clay_hardpan,per_muck,per_artificial,pool_metric_total,pool_depth_.cm.,bankfull_metric_total,average_bankfull_.m.,sinuosity,base_flow,date_last_rain,amt_last_rain,X._canopy_open,temp,dissolved_oxygen_mgl,ph,conductivity,fish_observed,fish_time_collecting_minutes,frogs_tadpoles_observed,salamanders_observed,salamander_time_collecting_minutes,aquatic_macros_observed,macro_time_collecting_minutes,mountain_dusky_larvae,mountain_dusky_juveniles,mountain_dusky_adults,northern_dusky_larvae,northern_dusky_juveniles,northern_dusky_adults,two_lined_larvae,two_lined_juveniles,two_lined_adults,long_tailed_larvae,long_tailed_juveniles,long_tailed_adults,northern_red_larvae,northern_red_juveniles,northern_red_adults,mole_larvae,mole_juveniles,mole_adults,four_toed_larvae,four_toed_juveniles,four_toed_adults,red_backed_juveniles,red_backed_adults,slimy_juveniles,slimy_adults,blacknose_dace,bluegill_sunfish,bluntnose_minnow,brindled_madtom,brook_stickleback,brook_trout,central_mudminnow,central_stoneroller_minnow,common_shiner,creek_chub,eastern_sand_darter,fantail_darter,fathead_minnow,golden_shiner,goldfish,grass_pickerel,greenside_darter,green_sunfish,johnny_darter,largemouth_bass,longear_sunfish,longnose_dace,mottled_sculpin,northern_hog_sucker,pumpkinseed_sunfish,rainbow_darter,rainbow_trout,redbelly_dace,redside_dace,river_chub,silverjaw_minnow,smallmouth_bass,spotfin_shiner,stonecat_madtom,striped_shiner,trout_perch,warmouth_sunfish,white_sucker,yellow_bullhead_catfish,unidentified_fish_fry,sessile_animals,aquatic_worms,sow_bugs,scuds,water_mites,damselfly_nymphs,alderfly_larvae,other_beetles,crayfish,dragonfly_nymphs,riffle_beetles,other_flies,midges,snails,clams,fishfly_larvae,water_penny_beetles,cranefly_larvae,ameletidae,arthropleidae,baetidae,baetiscidae,caenidae,ephemerellidae,ephemeridae,heptageniidae,isonychiidae,leptohyphidae,leptophlebiidae,polymitarcyidae,potamanthidae,pseudironidae,siphlonuridae,capniidae,chloroperlidae,leuctridae,nemouridae,peltoperlidae,perlidae,perlodidae,pteronarcyidae,taeniopterygidae,brachycentridae,dipseuodopsidae,glossosomatidae,goeridae,helicopsychidae,hydropsychidae,hydroptilidae,lepidostomatidae,leptoceridae,limnephilidae,molannidae,odontoceridae,philopotamidae,phryganeidae,polycentropodidae,psychomiidae,rhyacophilidae,uenoidae,total_mayfly,total_stonefly,total_caddisfly,total_ept,hhei_score,hmfei_score,hmfei_narrative) %>% distinct(stream_name, date, .keep_all = TRUE)

phwh_abiotic<- phwh_subset %>%
  dplyr::select(stream_name,date,substrate_metric_total,percent_best_types,per_boulder_slabs,per_boulder_256mm,per_bedrock,per_cobble,per_gravel,per_sand,per_silt,per_leaf_woody,per_detritus,per_clay_hardpan,per_muck,per_artificial,pool_metric_total,pool_depth_.cm.,bankfull_metric_total,average_bankfull_.m.,sinuosity,base_flow,date_last_rain,amt_last_rain,X._canopy_open,temp,dissolved_oxygen_mgl,ph,conductivity) %>% distinct(stream_name, date, .keep_all = TRUE)

phwh_salamander<- phwh_subset %>%
  dplyr::select(stream_name,date,mountain_dusky_larvae,mountain_dusky_juveniles,mountain_dusky_adults,northern_dusky_larvae,northern_dusky_juveniles,northern_dusky_adults,two_lined_larvae,two_lined_juveniles,two_lined_adults,long_tailed_larvae,long_tailed_juveniles,long_tailed_adults,northern_red_larvae,northern_red_juveniles,northern_red_adults,mole_larvae,mole_juveniles,mole_adults,four_toed_larvae,four_toed_juveniles,four_toed_adults,red_backed_juveniles,red_backed_adults,slimy_juveniles,slimy_adults) %>%  replace(is.na(.),0) %>% replace(.=="abundant",30) %>% replace(.=="common",6) %>% replace(.=="rare", 1) %>% replace(.=="very abundant",50) %>% 
  distinct(stream_name, date, .keep_all = TRUE)

phwh_fish<- phwh_subset %>%
  dplyr::select(stream_name,date,blacknose_dace,bluegill_sunfish,bluntnose_minnow,brindled_madtom,brook_stickleback,brook_trout,central_mudminnow,central_stoneroller_minnow,common_shiner,creek_chub,eastern_sand_darter,fantail_darter,fathead_minnow,golden_shiner,goldfish,grass_pickerel,greenside_darter,green_sunfish,johnny_darter,largemouth_bass,longear_sunfish,longnose_dace,mottled_sculpin,northern_hog_sucker,pumpkinseed_sunfish,rainbow_darter,rainbow_trout,redbelly_dace,redside_dace,river_chub,silverjaw_minnow,smallmouth_bass,spotfin_shiner,stonecat_madtom,striped_shiner,trout_perch,warmouth_sunfish,white_sucker,yellow_bullhead_catfish,unidentified_fish_fry) %>%
  replace(is.na(.),0) %>%  replace(is.na(.),0) %>% replace(.=="abundant",30) %>% replace(.=="common",6) %>% replace(.=="rare", 1) %>% replace(.=="very abundant",50) %>%
  distinct(stream_name, date, .keep_all = TRUE) %>% dplyr::select(-stream_name,-date)

phwh_fish<-sapply(phwh_fish,as.numeric)



phwh_macro<- phwh_subset %>%
  dplyr::select(stream_name,date,sessile_animals,aquatic_worms,sow_bugs,scuds,water_mites,damselfly_nymphs,alderfly_larvae,other_beetles,crayfish,dragonfly_nymphs,riffle_beetles,other_flies,midges,snails,clams,fishfly_larvae,water_penny_beetles,cranefly_larvae,total_mayfly,total_stonefly,total_caddisfly) %>%  replace(is.na(.),0) %>% replace(.=="abundant",30) %>% replace(.=="common",6) %>% replace(.=="very abundant",50) %>%
  replace(.=="rare", 1)  %>% distinct(stream_name, date, .keep_all = TRUE) %>% dplyr::select(-stream_name,-date)

phwh_macro<-sapply(phwh_macro,as.numeric)

phwh_ept<-phwh_subset %>%
  dplyr::select(stream_name,date,ameletidae,arthropleidae,baetidae,baetiscidae,caenidae,ephemerellidae,ephemeridae,heptageniidae,isonychiidae,leptohyphidae,leptophlebiidae,polymitarcyidae,potamanthidae,pseudironidae,siphlonuridae,capniidae,chloroperlidae,leuctridae,nemouridae,peltoperlidae,perlidae,perlodidae,pteronarcyidae,taeniopterygidae,brachycentridae,dipseuodopsidae,glossosomatidae,goeridae,helicopsychidae,hydropsychidae,hydroptilidae,lepidostomatidae,leptoceridae,limnephilidae,molannidae,odontoceridae,philopotamidae,phryganeidae,polycentropodidae,psychomiidae,rhyacophilidae,uenoidae) %>%  
  replace(is.na(.),0) %>% replace(.=="abundant",30) %>% replace(.=="common",6) %>% replace(.=="rare", 1)  %>% replace(.=="very abundant",50) %>%
  distinct(stream_name, date, .keep_all = TRUE)


phwh_e<-phwh_subset %>%
  dplyr::select(stream_name,date,ameletidae,arthropleidae,baetidae,baetiscidae,caenidae,ephemerellidae,ephemeridae,heptageniidae,isonychiidae,leptohyphidae,leptophlebiidae,polymitarcyidae,potamanthidae,pseudironidae,siphlonuridae) %>%  replace(is.na(.),0) %>% replace(.=="abundant",30) %>% replace(.=="common",6) %>% replace(.=="rare", 1)  %>% replace(.=="very abundant",50) %>%
  distinct(stream_name, date, .keep_all = TRUE)

phwh_p<-phwh_subset %>%
  dplyr::select(stream_name,date,capniidae,chloroperlidae,leuctridae,nemouridae,peltoperlidae,perlidae,perlodidae,pteronarcyidae,taeniopterygidae) %>%  
  replace(is.na(.),0) %>% replace(.=="abundant",30) %>% replace(.=="common",6) %>% replace(.=="rare", 1)  %>% replace(.=="very abundant",50) %>%
  distinct(stream_name, date, .keep_all = TRUE)

phwh_t<-phwh_subset %>%
  dplyr::select(stream_name,date,brachycentridae,dipseuodopsidae,glossosomatidae,goeridae,helicopsychidae,hydropsychidae,hydroptilidae,lepidostomatidae,leptoceridae,limnephilidae,molannidae,odontoceridae,philopotamidae,phryganeidae,polycentropodidae,psychomiidae,rhyacophilidae,uenoidae) %>%  
  replace(is.na(.),0) %>% replace(.=="abundant",30) %>% replace(.=="common",6) %>% replace(.=="rare", 1)  %>% replace(.=="very abundant",50) %>%
  distinct(stream_name, date, .keep_all = TRUE)

fish_sum<-phwh_subset$fish_observed
sala_sum<-phwh_subset$salamanders_observed
macro_sum<-phwh_subset$aquatic_macros_observed
frog_sum<-phwh_subset$frogs_tadpoles_observed
total_ept<-phwh_subset$total_ept
hhei<-phwh_subset$hhei_score
hmfei<-phwh_subset$hmfei_score
score<-phwh_subset$hmfei_narrative
reservation<-phwh_subset$reservation
stream_name<-phwh$stream_name
watershed<-phwh$watershed

reservationsum<-phwh_subset %>%
  distinct(stream_name, .keep_all = TRUE) %>%
  dplyr::select(stream_name,reservation) %>%
  arrange(., stream_name)

reservationsum<-na.omit(reservationsum)

reservationcol<-as.numeric(factor(reservationsum$reservation))



scoresum<-phwh_subset %>%
   distinct(stream_name, .keep_all = TRUE) %>%
  dplyr::select(stream_name,hmfei_narrative) %>%
  arrange(., stream_name)
scoresum<-na.omit(scoresum)

scorecol<-as.numeric(factor(scoresum$hmfei_narrative))

streambiotic<-cbind(phwh_salamander,phwh_fish,phwh_macro)
streambioticsum<-streambiotic %>% 
  group_by(stream_name) %>%
  dplyr::select(!date) %>%
  summarise_all(list(sum))


streambiotic_nosite<-streambioticsum[-1,-1] %>% 
  select_if(colSums(.) != 0)

taxatype<-c(rep("salamander",19),rep("fish",20),rep("macro",21))

```

# select sites 

```{r, warning = FALSE, message= FALSE}

sitechecks_surveyhistory_rename<-sitechecks_surveyhistory %>% rename(stream_name=stream_names)

sitechecks_surveyhistory_res<-merge(sitechecks_surveyhistory_rename,reservationsum,by="stream_name")

sitechecks_surveyhistory_ressum<-sitechecks_surveyhistory_res %>%
  group_by(reservation)

greenlightsites<-read.csv("G:/NaturalResources/Byer/DataAnalysis/PHWH/PHWH_cycle3sites_011323_sites.csv")

sitechecks_surveyhistory_res_suff<-sitechecks_surveyhistory_res[sitechecks_surveyhistory_res$stream_name %in% greenlightsites$stream_nam , ]


phwh_subset_suff<-phwh_subset[phwh_subset$stream_name %in% greenlightsites$stream_nam ,]


```


# just salamanders

```{r, warning = FALSE, message= FALSE}
class(phwh_subset_suff$stream_name)<-"character"

phwh_subset_suff_env_detecthistory<-phwh_subset_suff %>% 
  dplyr::select(stream_name,reservation,river_basin,latitude,longitude,date,substrate_metric_total,percent_best_types,per_boulder_slabs,per_boulder_256mm,
         per_bedrock,per_cobble,per_gravel,per_sand,per_silt,per_leaf_woody,per_detritus,per_clay_hardpan,per_muck,per_artificial,pool_metric_total,
         pool_depth_.cm.,bankfull_metric_total,average_bankfull_.m.,sinuosity,X._canopy_open,temp,
         dissolved_oxygen_mgl,ph,conductivity,hmfei_narrative) %>%
  group_by(stream_name, reservation,date)

phwh_salamander_suff<-phwh_salamander[phwh_salamander$stream_name %in% greenlightsites$stream_nam ,]


phwh_salamander_suff_detecthistory<-phwh_salamander_suff %>% 
  group_by(stream_name, date)

mtdusky<-phwh_salamander_suff_detecthistory$mountain_dusky_larvae+
  phwh_salamander_suff_detecthistory$mountain_dusky_juveniles+
  phwh_salamander_suff_detecthistory$mountain_dusky_adults

nodusky<-phwh_salamander_suff_detecthistory$northern_dusky_larvae+
  phwh_salamander_suff_detecthistory$northern_dusky_juveniles+
  phwh_salamander_suff_detecthistory$northern_dusky_adults

twolined<-phwh_salamander_suff_detecthistory$two_lined_larvae+
  phwh_salamander_suff_detecthistory$two_lined_juveniles+
  phwh_salamander_suff_detecthistory$two_lined_adults

longtail<-phwh_salamander_suff_detecthistory$long_tailed_larvae+
  phwh_salamander_suff_detecthistory$long_tailed_juveniles+
  phwh_salamander_suff_detecthistory$long_tailed_adults

red<-phwh_salamander_suff_detecthistory$northern_red_larvae+
  phwh_salamander_suff_detecthistory$northern_red_juveniles+
  phwh_salamander_suff_detecthistory$northern_red_adults

mole<-phwh_salamander_suff_detecthistory$mole_larvae+
  phwh_salamander_suff_detecthistory$mole_juveniles+
  phwh_salamander_suff_detecthistory$mole_adults

fourtoed<-phwh_salamander_suff_detecthistory$four_toed_larvae+
  phwh_salamander_suff_detecthistory$four_toed_juveniles+
  phwh_salamander_suff_detecthistory$four_toed_adults

redback<-phwh_salamander_suff_detecthistory$red_backed_juveniles+
  phwh_salamander_suff_detecthistory$red_backed_adults

slimy<-phwh_salamander_suff_detecthistory$slimy_juveniles+
  phwh_salamander_suff_detecthistory$slimy_adults

stream_name<-phwh_salamander_suff_detecthistory$stream_name
date<-phwh_salamander_suff_detecthistory$date

phwh_salamander_suff_sum<-data.frame(stream_name,date,
                                mtdusky,nodusky,twolined,longtail,red,
                                mole,fourtoed,redback,slimy)


sitechecks<-phwh_salamander_suff_sum[,c(1,2)]

sitechecks$date<- as.Date(sitechecks$date , format = "%Y/%m/%d")

sitechecks$dayofyear<-format(sitechecks$date, "%j")

sitechecks$datecount<-as.numeric(sitechecks$date-min(sitechecks$date)+1)


sitechecks_surveyhistory<-sitechecks %>%
  group_by(stream_name) %>% 
  mutate(check=as.character(dense_rank(date))) %>%
  arrange(stream_name) %>% 
  distinct(stream_name, datecount, .keep_all = TRUE) %>%
  spread(check,datecount) %>%
  summarise(`1` = mean(`1`,na.rm=T),
            `2`=mean(`2`,na.rm=T),
            `3`=mean(`3`,na.rm=T),
            `4`=mean(`4`,na.rm=T),
            `5`=mean(`5`,na.rm=T),
            `6`=mean(`6`,na.rm=T),
            `7`=mean(`7`,na.rm=T),
            `8`=mean(`8`,na.rm=T),
            `9` = mean(`9`,na.rm=T),
            `10`=mean(`10`,na.rm=T),
            `11`=mean(`11`,na.rm=T),
            `12`=mean(`12`,na.rm=T),
            `13`=mean(`13`,na.rm=T),
            `14`=mean(`14`,na.rm=T),
            `15`=mean(`15`,na.rm=T),
            `16`=mean(`16`,na.rm=T))
```

# all taxa

```{r, warning = FALSE, message= FALSE}
phwh_subset_biotic<-phwh %>%
  dplyr::select(stream_name,date,mountain_dusky_larvae,mountain_dusky_juveniles,
                mountain_dusky_adults,northern_dusky_larvae,northern_dusky_juveniles,northern_dusky_adults,two_lined_larvae,two_lined_juveniles,
                two_lined_adults,long_tailed_larvae,long_tailed_juveniles,long_tailed_adults,northern_red_larvae,northern_red_juveniles,northern_red_adults,
                mole_larvae,mole_juveniles,mole_adults,four_toed_larvae,four_toed_juveniles,four_toed_adults,red_backed_juveniles,red_backed_adults,
                slimy_juveniles,slimy_adults,blacknose_dace,bluegill_sunfish,bluntnose_minnow,brindled_madtom,brook_stickleback,brook_trout,
                central_mudminnow,central_stoneroller_minnow,common_shiner,creek_chub,eastern_sand_darter,fantail_darter,fathead_minnow,golden_shiner,
                goldfish,grass_pickerel,greenside_darter,green_sunfish,johnny_darter,largemouth_bass,longear_sunfish,longnose_dace,mottled_sculpin,
                northern_hog_sucker,pumpkinseed_sunfish,rainbow_darter,rainbow_trout,redbelly_dace,redside_dace,river_chub,silverjaw_minnow,smallmouth_bass,
                spotfin_shiner,stonecat_madtom,striped_shiner,trout_perch,warmouth_sunfish,white_sucker,yellow_bullhead_catfish,unidentified_fish_fry,
                sessile_animals,aquatic_worms,sow_bugs,scuds,water_mites,damselfly_nymphs,alderfly_larvae,other_beetles,crayfish,dragonfly_nymphs,
                riffle_beetles,other_flies,midges,snails,clams,fishfly_larvae,water_penny_beetles,cranefly_larvae,ameletidae,arthropleidae,baetidae,
                baetiscidae,caenidae,ephemerellidae,ephemeridae,heptageniidae,isonychiidae,leptohyphidae,leptophlebiidae,polymitarcyidae,potamanthidae,
                pseudironidae,siphlonuridae,capniidae,chloroperlidae,leuctridae,nemouridae,peltoperlidae,perlidae,perlodidae,pteronarcyidae,taeniopterygidae,
                brachycentridae,dipseuodopsidae,glossosomatidae,goeridae,helicopsychidae,hydropsychidae,hydroptilidae,lepidostomatidae,leptoceridae,
                limnephilidae,molannidae,odontoceridae,philopotamidae,phryganeidae,polycentropodidae,psychomiidae,rhyacophilidae,uenoidae) %>% 
  distinct(stream_name, date, .keep_all = TRUE)

phwh_subset_biotic_2<-phwh_subset_biotic %>% replace(is.na(.),0) %>% replace(.=="abundant",30) %>% replace(.=="common",6) %>% replace(.=="rare", 1) %>% replace(.=="very abundant",50) %>%
  distinct(stream_name, date, .keep_all = TRUE) %>% dplyr::select(-stream_name,-date)

phwh_subset_biotic_3<-data.frame(sapply(phwh_subset_biotic_2,as.numeric))

phwh_subset_biotic_3$stream_name<-phwh_subset_biotic$stream_name

phwh_subset_biotic_3$date<-phwh_subset_biotic$date

streambiotic<-phwh_subset_biotic_3
streambioticsum<-streambiotic %>% 
  group_by(stream_name) %>%
  dplyr::select(!date) %>%
  summarise_all(list(sum))

streambioticsum_suff <-streambioticsum[streambioticsum$stream_name %in% greenlightsites$stream_nam , ]

streambioticsum_suff_mtdusky<-streambioticsum_suff$mountain_dusky_larvae+
  streambioticsum_suff$mountain_dusky_juveniles+
  streambioticsum_suff$mountain_dusky_adults

streambioticsum_suff_nodusky<-streambioticsum_suff$northern_dusky_larvae+
  streambioticsum_suff$northern_dusky_juveniles+
  streambioticsum_suff$northern_dusky_adults

streambioticsum_suff_twolined<-streambioticsum_suff$two_lined_larvae+
  streambioticsum_suff$two_lined_juveniles+
  streambioticsum_suff$two_lined_adults

streambioticsum_suff_longtail<-streambioticsum_suff$long_tailed_larvae+
  streambioticsum_suff$long_tailed_juveniles+
  streambioticsum_suff$long_tailed_adults

streambioticsum_suff_red<-streambioticsum_suff$northern_red_larvae+
  streambioticsum_suff$northern_red_juveniles+
  streambioticsum_suff$northern_red_adults

streambioticsum_suff_mole<-streambioticsum_suff$mole_larvae+
  streambioticsum_suff$mole_juveniles+
  streambioticsum_suff$mole_adults

streambioticsum_suff_fourtoed<-streambioticsum_suff$four_toed_larvae+
  streambioticsum_suff$four_toed_juveniles+
  streambioticsum_suff$four_toed_adults

streambioticsum_suff_redback<-streambioticsum_suff$red_backed_juveniles+
  streambioticsum_suff$red_backed_adults

streambioticsum_suff_slimy<-streambioticsum_suff$slimy_juveniles+
  streambioticsum_suff$slimy_adults

streambioticsum_suff_sal<-data.frame(streambioticsum_suff_mtdusky,streambioticsum_suff_nodusky,streambioticsum_suff_twolined,
                                     streambioticsum_suff_longtail,streambioticsum_suff_red,streambioticsum_suff_mole,
                                     streambioticsum_suff_fourtoed,streambioticsum_suff_redback,streambioticsum_suff_slimy)

streambioticsum_suff_fish<-streambioticsum_suff[,27:65]

streambioticsum_suff_macro<-streambioticsum_suff[,67:126]

streambioticsum_suff_biotic<-data.frame(streambioticsum_suff_sal,streambioticsum_suff_fish,streambioticsum_suff_macro)

streambioticsum_suff_biotic_bin<- streambioticsum_suff_biotic %>% mutate_if(is.numeric, ~1 * (. >= 1))
```


# make new dataset for occupancy

```{r, warning = FALSE, message= FALSE}

streambiotic_suff<-phwh_subset_biotic_3[phwh_subset_biotic_3$stream_name %in% greenlightsites$stream_nam , ]

streambiotic_suff_mtdusky<-streambiotic_suff$mountain_dusky_larvae+
  streambiotic_suff$mountain_dusky_juveniles+
  streambiotic_suff$mountain_dusky_adults

streambiotic_suff_nodusky<-streambiotic_suff$northern_dusky_larvae+
  streambiotic_suff$northern_dusky_juveniles+
  streambiotic_suff$northern_dusky_adults

streambiotic_suff_twolined<-streambiotic_suff$two_lined_larvae+
  streambiotic_suff$two_lined_juveniles+
  streambiotic_suff$two_lined_adults

streambiotic_suff_longtail<-streambiotic_suff$long_tailed_larvae+
  streambiotic_suff$long_tailed_juveniles+
  streambiotic_suff$long_tailed_adults

streambiotic_suff_red<-streambiotic_suff$northern_red_larvae+
  streambiotic_suff$northern_red_juveniles+
  streambiotic_suff$northern_red_adults

streambiotic_suff_mole<-streambiotic_suff$mole_larvae+
  streambiotic_suff$mole_juveniles+
  streambiotic_suff$mole_adults

streambiotic_suff_fourtoed<-streambiotic_suff$four_toed_larvae+
  streambiotic_suff$four_toed_juveniles+
  streambiotic_suff$four_toed_adults

streambiotic_suff_redback<-streambiotic_suff$red_backed_juveniles+
  streambiotic_suff$red_backed_adults

streambiotic_suff_slimy<-streambiotic_suff$slimy_juveniles+
  streambiotic_suff$slimy_adults

streambiotic_suff_sal<-data.frame(streambiotic_suff_mtdusky,streambiotic_suff_nodusky,streambiotic_suff_twolined,
                                  streambiotic_suff_longtail,streambiotic_suff_red,streambiotic_suff_mole,
                                  streambiotic_suff_fourtoed,streambiotic_suff_redback,streambiotic_suff_slimy)

streambiotic_suff_fish<-streambiotic_suff[,26:64]

streambiotic_suff_macro<-streambiotic_suff[,66:125]

streambiotic_suff_biotic<-data.frame(streambiotic_suff_sal,streambiotic_suff_fish,streambiotic_suff_macro)

streambiotic_suff_biotic_bin<- streambiotic_suff_biotic %>% mutate_if(is.numeric, ~1 * (. >= 1))

phwh_alltax_suff_env_detecthistory<-cbind(phwh_subset_suff_env_detecthistory,streambiotic_suff_biotic_bin)

phwh_alltax_suff_env_detecthistory$date<- as.Date(phwh_alltax_suff_env_detecthistory$date , format = "%Y/%m/%d")

phwh_alltax_suff_env_detecthistory$datecount<-as.numeric(phwh_alltax_suff_env_detecthistory$date-min(phwh_alltax_suff_env_detecthistory$date)+1)

phwh_alltax_suff_env_detecthistory$dayofyear<-format(phwh_alltax_suff_env_detecthistory$date, "%j")



phwh_alltax_suff_env_detecthistory<-phwh_alltax_suff_env_detecthistory %>%
  dplyr::select(!c(latitude,longitude))


phwh_alltax_suff_env_detecthistory_firstcheck<- phwh_alltax_suff_env_detecthistory %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[1]))) %>% 
  unnest(cols = c(reservation, river_basin, date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

firstcheck_all<-phwh_alltax_suff_env_detecthistory_firstcheck %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")

phwh_alltax_suff_env_detecthistory_secondcheck<- phwh_alltax_suff_env_detecthistory %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[2]))) %>% 
  unnest(cols = c(reservation, river_basin, date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

secondcheck_all<-phwh_alltax_suff_env_detecthistory_secondcheck %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")


phwh_alltax_suff_env_detecthistory_thirdcheck<- phwh_alltax_suff_env_detecthistory %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[3]))) %>% 
  unnest(cols = c(reservation, river_basin, date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

thirdcheck_all<-phwh_alltax_suff_env_detecthistory_thirdcheck %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")

phwh_alltax_suff_env_detecthistory_fourthcheck<- phwh_alltax_suff_env_detecthistory %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[4]))) %>% 
  unnest(cols = c(reservation, river_basin, date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

fourthcheck_all<-phwh_alltax_suff_env_detecthistory_fourthcheck %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")

phwh_alltax_suff_env_detecthistory_fifthcheck<- phwh_alltax_suff_env_detecthistory %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[5]))) %>% 
  unnest(cols = c(reservation, river_basin, date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

fifthcheck_all<-phwh_alltax_suff_env_detecthistory_fifthcheck %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")

phwh_alltax_suff_env_detecthistory_sixthcheck<- phwh_alltax_suff_env_detecthistory %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[6]))) %>% 
  unnest(cols = c(reservation, river_basin, date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

sixthcheck_all<-phwh_alltax_suff_env_detecthistory_sixthcheck %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")

phwh_alltax_suff_env_detecthistory_seventhcheck<- phwh_alltax_suff_env_detecthistory %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[7]))) %>% 
  unnest(cols = c(reservation, river_basin, date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

seventhcheck_all<-phwh_alltax_suff_env_detecthistory_seventhcheck %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")

seventhcheck_all_2<-phwh_alltax_suff_env_detecthistory_seventhcheck %>%
  dplyr::select(stream_name,streambiotic_suff_mtdusky:uenoidae) %>%
  pivot_longer(-stream_name) %>%
  pivot_wider(names_from=stream_name) %>%
  column_to_rownames(var = "name")

phwh_alltax_suff_env_detecthistory_eighthcheck<- phwh_alltax_suff_env_detecthistory %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[8]))) %>% 
  unnest(cols = c(reservation, river_basin, date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

eighthcheck_all<-phwh_alltax_suff_env_detecthistory_eighthcheck %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")




phwhall_occobject<-list()

phwhall_occobject$y<-array(dim=c(108,137,8))

phwhall_occobject$y <- array(c(as.matrix(firstcheck_all),as.matrix(secondcheck_all),
                               as.matrix(thirdcheck_all),as.matrix(fourthcheck_all),
                               as.matrix(fifthcheck_all),as.matrix(sixthcheck_all),
                               as.matrix(seventhcheck_all),as.matrix(eighthcheck_all)), dim=c(108,137,8))

phwhall_occobject$det.covs$day<-as.matrix(sitechecks_surveyhistory[,2:9])

seasons<-round(sitechecks_surveyhistory[,2:9]/365)+1

dayyear<-cbind(as.numeric(phwh_alltax_suff_env_detecthistory_firstcheck$dayofyear),as.numeric(phwh_alltax_suff_env_detecthistory_secondcheck$dayofyear),
               as.numeric(phwh_alltax_suff_env_detecthistory_thirdcheck$dayofyear),as.numeric(phwh_alltax_suff_env_detecthistory_fourthcheck$dayofyear),
               as.numeric(phwh_alltax_suff_env_detecthistory_fifthcheck$dayofyear),as.numeric(phwh_alltax_suff_env_detecthistory_sixthcheck$dayofyear),
               as.numeric(phwh_alltax_suff_env_detecthistory_seventhcheck$dayofyear),as.numeric(phwh_alltax_suff_env_detecthistory_eighthcheck$dayofyear))


phwhall_occobject$det.covs$seasons<-as.matrix(seasons)

phwhall_occobject$det.covs$dayyear<-as.matrix(dayyear)

occcov<-phwh_subset_suff_env_detecthistory %>% group_by(stream_name) %>% 
  summarise_if(is.numeric, mean, na.rm = TRUE)

occcov_2<-occcov[,-1:-4]

cols<-seq(1,23,1)

occcov_2[,cols] = apply(occcov_2[,cols], 2, function(x) as.numeric(as.character(x)))

occcov_3<-occcov_2[ , colSums(is.na(occcov_2)) == 0]

phwhall_occobject$occ.covs<-as.matrix(occcov_3)


library(rgdal)
d <- data.frame(lon=occcov$longitude, lat=occcov$latitude)
coordinates(d) <- c("lon", "lat")
proj4string(d) <- CRS("+init=epsg:4326") # WGS 84
CRS.new <- CRS("+init=epsg:3734")
d.3734 <- spTransform(d, CRS.new)
plot(d)
projcoords<-data.frame(d.3734)



phwhall_occobject$coords<-as.matrix(projcoords)

phwhall_occobject$occ.covs


rownames(phwhall_occobject$y)<-rownames(firstcheck_all)

```

## single year of surveying?

```{r,warning=FALSE,message=FALSE}

phwh_alltax_suff_env_detecthistory<-cbind(phwh_subset_suff_env_detecthistory,streambiotic_suff_biotic_bin)

phwh_alltax_suff_env_detecthistory$date<- as.Date(phwh_alltax_suff_env_detecthistory$date , format = "%Y/%m/%d")

phwh_alltax_suff_env_detecthistory$datecount<-as.numeric(phwh_alltax_suff_env_detecthistory$date-min(phwh_alltax_suff_env_detecthistory$date)+1)

phwh_alltax_suff_env_detecthistory$dayofyear<-format(phwh_alltax_suff_env_detecthistory$date, "%j")


phwh_alltax_suff_env_detecthistory_years<-phwh_alltax_suff_env_detecthistory

phwh_alltax_suff_env_detecthistory_years$year<-year(phwh_alltax_suff_env_detecthistory$date)

phwh_alltax_suff_env_detecthistory_years_2 <- phwh_alltax_suff_env_detecthistory_years %>%
  group_by(stream_name,year,dayofyear)

summary(factor(phwh_alltax_suff_env_detecthistory_years$year))

phwh_alltax_suff_env_detecthistory_years_2008<-phwh_alltax_suff_env_detecthistory_years[phwh_alltax_suff_env_detecthistory_years$year %in% 2008,] %>%
  dplyr::select(-year)


phwh_alltax_suff_env_detecthistory_years_2008_firstcheck<- phwh_alltax_suff_env_detecthistory_years_2008 %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[1]))) %>% 
  unnest(cols = c(reservation, river_basin, latitude,longitude, date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

firstcheck_all_2008<-phwh_alltax_suff_env_detecthistory_years_2008_firstcheck %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")

phwh_alltax_suff_env_detecthistory_secondcheck_2008<- phwh_alltax_suff_env_detecthistory_years_2008 %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[2]))) %>% 
  unnest(cols = c(reservation, river_basin, latitude,longitude,date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

secondcheck_all_2008<-phwh_alltax_suff_env_detecthistory_secondcheck_2008 %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")


phwh_alltax_suff_env_detecthistory_thirdcheck_2008<- phwh_alltax_suff_env_detecthistory_years_2008 %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[3]))) %>% 
  unnest(cols = c(reservation, river_basin, latitude,longitude,date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

thirdcheck_all_2008<-phwh_alltax_suff_env_detecthistory_thirdcheck_2008 %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")

#stop at third survey - nothing further


sitechecks_2008<-phwh_alltax_suff_env_detecthistory_years_2008[,c(1,6)]

sitechecks_2008$date<- as.Date(sitechecks_2008$date , format = "%Y/%m/%d")

sitechecks_2008$dayofyear<-format(sitechecks_2008$date, "%j")

sitechecks_2008$datecount<-as.numeric(sitechecks_2008$date-min(sitechecks_2008$date)+1)


sitechecks_2008_surveyhistory<-sitechecks_2008 %>%
  group_by(stream_name) %>% 
  mutate(check=as.character(dense_rank(date))) %>%
  arrange(stream_name) %>% 
  distinct(stream_name, datecount, .keep_all = TRUE) %>%
  spread(check,datecount) %>%
  summarise(`1` = mean(`1`,na.rm=T),
            `2`=mean(`2`,na.rm=T),
            `3`=mean(`3`,na.rm=T),
            )

phwhall_occobject_2008<-list()

phwhall_occobject_2008$y<-array(dim=c(108,36,3))

phwhall_occobject_2008$y <- array(c(as.matrix(firstcheck_all_2008),as.matrix(secondcheck_all_2008),
                               as.matrix(thirdcheck_all_2008)), dim=c(108,36,3))

phwhall_occobject_2008$det.covs$day<-as.matrix(sitechecks_2008_surveyhistory[,2:4])

dayyear_2008<-cbind(as.numeric(phwh_alltax_suff_env_detecthistory_years_2008_firstcheck$dayofyear),as.numeric(phwh_alltax_suff_env_detecthistory_secondcheck_2008$dayofyear),
               as.numeric(phwh_alltax_suff_env_detecthistory_thirdcheck_2008$dayofyear))

phwhall_occobject_2008$det.covs$dayyear<-as.matrix(dayyear_2008)

phwh_subset_suff_env_detecthistory_2008<-phwh_alltax_suff_env_detecthistory_years_2008 %>% 
  dplyr::select(stream_name,reservation,latitude,longitude,river_basin,date,substrate_metric_total,percent_best_types,per_boulder_slabs,per_boulder_256mm,
         per_bedrock,per_cobble,per_gravel,per_sand,per_silt,per_leaf_woody,per_detritus,per_clay_hardpan,per_muck,per_artificial,pool_metric_total,
         pool_depth_.cm.,bankfull_metric_total,average_bankfull_.m.,sinuosity,X._canopy_open,temp,
         dissolved_oxygen_mgl,ph,conductivity,hmfei_narrative) %>%
  group_by(stream_name, reservation,date)

occcov_2008<-phwh_subset_suff_env_detecthistory_2008 %>% group_by(stream_name) %>% 
  summarise_if(is.numeric, mean, na.rm = TRUE)

occcov_2008_2<-occcov_2008[,c(-1,-4)]

cols<-seq(1,25,1)

occcov_2008_2[,cols] = apply(occcov_2008_2[,cols], 2, function(x) as.numeric(as.character(x)))

occcov_2008_3<-occcov_2008_2[ , colSums(is.na(occcov_2008_2)) == 0]

phwhall_occobject_2008$occ.covs<-as.matrix(occcov_2008_3)

library(rgdal)
d <- data.frame(lon=occcov_2008_3$longitude, lat=occcov_2008_3$latitude)
coordinates(d) <- c("lon", "lat")
proj4string(d) <- CRS("+init=epsg:4326") # WGS 84
CRS.new <- CRS("+init=epsg:3734")
d.3734 <- spTransform(d, CRS.new)
plot(d)
projcoords<-data.frame(d.3734)

phwhall_occobject_2008$coords<-as.matrix(projcoords)

rownames(phwhall_occobject_2008$y)<-rownames(firstcheck_all_2008)



phwh_alltax_suff_env_detecthistory_years_2009<-phwh_alltax_suff_env_detecthistory_years[phwh_alltax_suff_env_detecthistory_years$year %in% 2009,]

phwh_alltax_suff_env_detecthistory_years_2014<-phwh_alltax_suff_env_detecthistory_years[phwh_alltax_suff_env_detecthistory_years$year %in% 2014,]

phwh_alltax_suff_env_detecthistory_years_2015<-phwh_alltax_suff_env_detecthistory_years[phwh_alltax_suff_env_detecthistory_years$year %in% 2015,]

phwh_alltax_suff_env_detecthistory_years_2018<-phwh_alltax_suff_env_detecthistory_years[phwh_alltax_suff_env_detecthistory_years$year %in% 2018,]

occcov_4<-occcov_3
occcov_4$stream_name<-occcov$stream_name

write.csv(occcov_4,"PHWH_occcovs_mean.csv")
```

#First cycle surveys - 2007-2010

```{r,warning=FALSE,message=FALSE}

phwh_alltax_suff_env_detecthistory_years_0710<-phwh_alltax_suff_env_detecthistory_years[phwh_alltax_suff_env_detecthistory_years$year %in% c(2007,2008,2009,2010),] %>%
  dplyr::select(-year)


phwh_alltax_suff_env_detecthistory_years_0710_firstcheck<- phwh_alltax_suff_env_detecthistory_years_0710 %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[1]))) %>% 
  unnest(cols = c(reservation, river_basin, latitude,longitude, date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

firstcheck_all_0710<-phwh_alltax_suff_env_detecthistory_years_0710_firstcheck %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")

phwh_alltax_suff_env_detecthistory_secondcheck_0710<- phwh_alltax_suff_env_detecthistory_years_0710 %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[2]))) %>% 
  unnest(cols = c(reservation, river_basin, latitude,longitude,date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

secondcheck_all_0710<-phwh_alltax_suff_env_detecthistory_secondcheck_0710 %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")


phwh_alltax_suff_env_detecthistory_thirdcheck_0710<- phwh_alltax_suff_env_detecthistory_years_0710 %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[3]))) %>% 
  unnest(cols = c(reservation, river_basin, latitude,longitude,date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

thirdcheck_all_0710<-phwh_alltax_suff_env_detecthistory_thirdcheck_0710 %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")

phwh_alltax_suff_env_detecthistory_fourthcheck_0710<- phwh_alltax_suff_env_detecthistory_years_0710 %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[4]))) %>% 
  unnest(cols = c(reservation, river_basin, latitude,longitude,date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

fourthcheck_all_0710<-phwh_alltax_suff_env_detecthistory_fourthcheck_0710 %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")


phwh_alltax_suff_env_detecthistory_fifthcheck_0710<- phwh_alltax_suff_env_detecthistory_years_0710 %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[5]))) %>% 
  unnest(cols = c(reservation, river_basin, latitude,longitude,date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

fifthcheck_all_0710<-phwh_alltax_suff_env_detecthistory_fifthcheck_0710 %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")

#stop at third survey - nothing further


sitechecks_0710<-phwh_alltax_suff_env_detecthistory_years_0710[,c(1,6)]

sitechecks_0710$date<- as.Date(sitechecks_0710$date , format = "%Y/%m/%d")

sitechecks_0710$dayofyear<-format(sitechecks_0710$date, "%j")

sitechecks_0710$datecount<-as.numeric(sitechecks_0710$date-min(sitechecks_0710$date)+1)


sitechecks_0710_surveyhistory<-sitechecks_0710 %>%
  group_by(stream_name) %>% 
  mutate(check=as.character(dense_rank(date))) %>%
  arrange(stream_name) %>% 
  distinct(stream_name, datecount, .keep_all = TRUE) %>%
  spread(check,datecount) %>%
  summarise(`1` = mean(`1`,na.rm=T),
            `2`=mean(`2`,na.rm=T),
            `3`=mean(`3`,na.rm=T),
            )

phwhall_occobject_0710<-list()

phwhall_occobject_0710$y<-array(dim=c(108,141,3))

phwhall_occobject_0710$y <- array(c(as.matrix(firstcheck_all_0710),as.matrix(secondcheck_all_0710),
                               as.matrix(thirdcheck_all_0710)), dim=c(108,141,3))

phwhall_occobject_0710$det.covs$day<-as.matrix(sitechecks_0710_surveyhistory[,2:4])

dayyear_0710<-cbind(as.numeric(phwh_alltax_suff_env_detecthistory_years_0710_firstcheck$dayofyear),as.numeric(phwh_alltax_suff_env_detecthistory_secondcheck_0710$dayofyear),
               as.numeric(phwh_alltax_suff_env_detecthistory_thirdcheck_0710$dayofyear))

phwhall_occobject_0710$det.covs$dayyear<-as.matrix(dayyear_0710)

phwh_subset_suff_env_detecthistory_0710<-phwh_alltax_suff_env_detecthistory_years_0710 %>% 
  dplyr::select(stream_name,reservation,latitude,longitude,river_basin,date,substrate_metric_total,percent_best_types,per_boulder_slabs,per_boulder_256mm,
         per_bedrock,per_cobble,per_gravel,per_sand,per_silt,per_leaf_woody,per_detritus,per_clay_hardpan,per_muck,per_artificial,pool_metric_total,
         pool_depth_.cm.,bankfull_metric_total,average_bankfull_.m.,sinuosity,X._canopy_open,temp,
         dissolved_oxygen_mgl,ph,conductivity,hmfei_narrative) %>%
  group_by(stream_name, reservation,date)

occcov_0710<-phwh_subset_suff_env_detecthistory_0710 %>% group_by(stream_name) %>% 
  summarise_if(is.numeric, mean, na.rm = TRUE)

occcov_0710_2<-occcov_0710[,c(-1,-4)]

cols<-seq(1,25,1)

occcov_0710_2[,cols] = apply(occcov_0710_2[,cols], 2, function(x) as.numeric(as.character(x)))

occcov_0710_3<-occcov_0710_2[ , colSums(is.na(occcov_0710_2)) == 0]

phwhall_occobject_0710$occ.covs<-as.matrix(occcov_0710_3)

library(rgdal)
d <- data.frame(lon=occcov_0710_3$longitude, lat=occcov_0710_3$latitude)
coordinates(d) <- c("lon", "lat")
proj4string(d) <- CRS("+init=epsg:4326") # WGS 84
CRS.new <- CRS("+init=epsg:3734")
d.3734 <- spTransform(d, CRS.new)
plot(d)
projcoords<-data.frame(d.3734)

phwhall_occobject_0710$coords<-as.matrix(projcoords)

rownames(phwhall_occobject_0710$y)<-rownames(firstcheck_all_0710)

```

#First cycle surveys - 2013-2016

```{r,warning=FALSE,message=FALSE}

phwh_alltax_suff_env_detecthistory_years_1316<-phwh_alltax_suff_env_detecthistory_years[phwh_alltax_suff_env_detecthistory_years$year %in% c(2013,2014,2015,2016),] %>%
  dplyr::select(-year)


phwh_alltax_suff_env_detecthistory_years_1316_firstcheck<- phwh_alltax_suff_env_detecthistory_years_1316 %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[1]))) %>% 
  unnest(cols = c(reservation, river_basin, latitude,longitude, date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

firstcheck_all_1316<-phwh_alltax_suff_env_detecthistory_years_1316_firstcheck %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")

phwh_alltax_suff_env_detecthistory_secondcheck_1316<- phwh_alltax_suff_env_detecthistory_years_1316 %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[2]))) %>% 
  unnest(cols = c(reservation, river_basin, latitude,longitude,date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

secondcheck_all_1316<-phwh_alltax_suff_env_detecthistory_secondcheck_1316 %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")


phwh_alltax_suff_env_detecthistory_thirdcheck_1316<- phwh_alltax_suff_env_detecthistory_years_1316 %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[3]))) %>% 
  unnest(cols = c(reservation, river_basin, latitude,longitude,date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

thirdcheck_all_1316<-phwh_alltax_suff_env_detecthistory_thirdcheck_1316 %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")

phwh_alltax_suff_env_detecthistory_fourthcheck_1316<- phwh_alltax_suff_env_detecthistory_years_1316 %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[4]))) %>% 
  unnest(cols = c(reservation, river_basin, latitude,longitude,date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

fourthcheck_all_1316<-phwh_alltax_suff_env_detecthistory_fourthcheck_1316 %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")


phwh_alltax_suff_env_detecthistory_fifthcheck_1316<- phwh_alltax_suff_env_detecthistory_years_1316 %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[5]))) %>% 
  unnest(cols = c(reservation, river_basin, latitude,longitude,date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

fifthcheck_all_1316<-phwh_alltax_suff_env_detecthistory_fifthcheck_1316 %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")

#stop at third survey - not much further


sitechecks_1316<-phwh_alltax_suff_env_detecthistory_years_1316[,c(1,6)]

sitechecks_1316$date<- as.Date(sitechecks_1316$date , format = "%Y/%m/%d")

sitechecks_1316$dayofyear<-format(sitechecks_1316$date, "%j")

sitechecks_1316$datecount<-as.numeric(sitechecks_1316$date-min(sitechecks_1316$date)+1)


sitechecks_1316_surveyhistory<-sitechecks_1316 %>%
  group_by(stream_name) %>% 
  mutate(check=as.character(dense_rank(date))) %>%
  arrange(stream_name) %>% 
  distinct(stream_name, datecount, .keep_all = TRUE) %>%
  spread(check,datecount) %>%
  summarise(`1` = mean(`1`,na.rm=T),
            `2`=mean(`2`,na.rm=T),
            `3`=mean(`3`,na.rm=T),
            )

phwhall_occobject_1316<-list()

phwhall_occobject_1316$y<-array(dim=c(108,154,3))

phwhall_occobject_1316$y <- array(c(as.matrix(firstcheck_all_1316),as.matrix(secondcheck_all_1316),
                               as.matrix(thirdcheck_all_1316)), dim=c(108,154,3))

phwhall_occobject_1316$det.covs$day<-as.matrix(sitechecks_1316_surveyhistory[,2:4])

dayyear_1316<-cbind(as.numeric(phwh_alltax_suff_env_detecthistory_years_1316_firstcheck$dayofyear),as.numeric(phwh_alltax_suff_env_detecthistory_secondcheck_1316$dayofyear),
               as.numeric(phwh_alltax_suff_env_detecthistory_thirdcheck_1316$dayofyear))

phwhall_occobject_1316$det.covs$dayyear<-as.matrix(dayyear_1316)

phwh_subset_suff_env_detecthistory_1316<-phwh_alltax_suff_env_detecthistory_years_1316 %>% 
  dplyr::select(stream_name,reservation,latitude,longitude,river_basin,date,substrate_metric_total,percent_best_types,per_boulder_slabs,per_boulder_256mm,
         per_bedrock,per_cobble,per_gravel,per_sand,per_silt,per_leaf_woody,per_detritus,per_clay_hardpan,per_muck,per_artificial,pool_metric_total,
         pool_depth_.cm.,bankfull_metric_total,average_bankfull_.m.,sinuosity,X._canopy_open,temp,
         dissolved_oxygen_mgl,ph,conductivity,hmfei_narrative) %>%
  group_by(stream_name, reservation,date)

occcov_1316<-phwh_subset_suff_env_detecthistory_1316 %>% group_by(stream_name) %>% 
  summarise_if(is.numeric, mean, na.rm = TRUE)

occcov_1316_2<-occcov_1316[,c(-1,-4)]

cols<-seq(1,25,1)

occcov_1316_2[,cols] = apply(occcov_1316_2[,cols], 2, function(x) as.numeric(as.character(x)))

occcov_1316_3<-occcov_1316_2[ , colSums(is.na(occcov_1316_2)) == 0]

phwhall_occobject_1316$occ.covs<-as.matrix(occcov_1316_3)

library(rgdal)
d <- data.frame(lon=occcov_1316_3$longitude, lat=occcov_1316_3$latitude)
coordinates(d) <- c("lon", "lat")
proj4string(d) <- CRS("+init=epsg:4326") # WGS 84
CRS.new <- CRS("+init=epsg:3734")
d.3734 <- spTransform(d, CRS.new)
plot(d)
projcoords<-data.frame(d.3734)

phwhall_occobject_1316$coords<-as.matrix(projcoords)

rownames(phwhall_occobject_1316$y)<-rownames(firstcheck_all_1316)

```

#First cycle surveys - 2017-2021

```{r,warning=FALSE,message=FALSE}

phwh_alltax_suff_env_detecthistory_years_1821<-phwh_alltax_suff_env_detecthistory_years[phwh_alltax_suff_env_detecthistory_years$year %in% c(2018,2019,2020,2021),] %>%
  dplyr::select(-year)


phwh_alltax_suff_env_detecthistory_years_1821_firstcheck<- phwh_alltax_suff_env_detecthistory_years_1821 %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[1]))) %>% 
  unnest(cols = c(reservation, river_basin, latitude,longitude, date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

firstcheck_all_1821<-phwh_alltax_suff_env_detecthistory_years_1821_firstcheck %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")

phwh_alltax_suff_env_detecthistory_secondcheck_1821<- phwh_alltax_suff_env_detecthistory_years_1821 %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[2]))) %>% 
  unnest(cols = c(reservation, river_basin, latitude,longitude,date, substrate_metric_total, percent_best_types, 
                  per_boulder_slabs, per_boulder_256mm, per_bedrock, per_cobble, 
                  per_gravel, per_sand, per_silt, per_leaf_woody, per_detritus, 
                  per_clay_hardpan, per_muck, per_artificial, pool_metric_total, 
                  pool_depth_.cm., bankfull_metric_total, average_bankfull_.m., 
                  sinuosity, X._canopy_open, temp, dissolved_oxygen_mgl, ph, 
                  conductivity, hmfei_narrative, streambiotic_suff_mtdusky, 
                  streambiotic_suff_nodusky, streambiotic_suff_twolined, streambiotic_suff_longtail, 
                  streambiotic_suff_red, streambiotic_suff_mole, streambiotic_suff_fourtoed, 
                  streambiotic_suff_redback, streambiotic_suff_slimy, blacknose_dace, 
                  bluegill_sunfish, bluntnose_minnow, brindled_madtom, brook_stickleback, 
                  brook_trout, central_mudminnow, central_stoneroller_minnow, 
                  common_shiner, creek_chub, eastern_sand_darter, fantail_darter, 
                  fathead_minnow, golden_shiner, goldfish, grass_pickerel, 
                  greenside_darter, green_sunfish, johnny_darter, largemouth_bass, 
                  longear_sunfish, longnose_dace, mottled_sculpin, northern_hog_sucker, 
                  pumpkinseed_sunfish, rainbow_darter, rainbow_trout, redbelly_dace, 
                  redside_dace, river_chub, silverjaw_minnow, smallmouth_bass, 
                  spotfin_shiner, stonecat_madtom, striped_shiner, trout_perch, 
                  warmouth_sunfish, white_sucker, yellow_bullhead_catfish, 
                  sessile_animals, aquatic_worms, sow_bugs, scuds, water_mites, 
                  damselfly_nymphs, alderfly_larvae, other_beetles, crayfish, 
                  dragonfly_nymphs, riffle_beetles, other_flies, midges, snails, 
                  clams, fishfly_larvae, water_penny_beetles, cranefly_larvae, 
                  ameletidae, arthropleidae, baetidae, baetiscidae, caenidae, 
                  ephemerellidae, ephemeridae, heptageniidae, isonychiidae, 
                  leptohyphidae, leptophlebiidae, polymitarcyidae, potamanthidae, 
                  pseudironidae, siphlonuridae, capniidae, chloroperlidae, 
                  leuctridae, nemouridae, peltoperlidae, perlidae, perlodidae, 
                  pteronarcyidae, taeniopterygidae, brachycentridae, dipseuodopsidae, 
                  glossosomatidae, goeridae, helicopsychidae, hydropsychidae, 
                  hydroptilidae, lepidostomatidae, leptoceridae, limnephilidae, 
                  molannidae, odontoceridae, philopotamidae, phryganeidae, 
                  polycentropodidae, psychomiidae, rhyacophilidae, uenoidae, 
                  datecount, dayofyear))

secondcheck_all_1821<-phwh_alltax_suff_env_detecthistory_secondcheck_1821 %>%
  dplyr::select(streambiotic_suff_mtdusky:uenoidae) %>%
  rownames_to_column() %>%
  pivot_longer(-rowname) %>%
  pivot_wider(names_from=rowname) %>%
  column_to_rownames(var = "name")



#stop at second survey - not much further

sitechecks_1821<-phwh_alltax_suff_env_detecthistory_years_1821[,c(1,6)]

sitechecks_1821$date<- as.Date(sitechecks_1821$date , format = "%Y/%m/%d")

sitechecks_1821$dayofyear<-format(sitechecks_1821$date, "%j")

sitechecks_1821$datecount<-as.numeric(sitechecks_1821$date-min(sitechecks_1821$date)+1)


sitechecks_1821_surveyhistory<-sitechecks_1821 %>%
  group_by(stream_name) %>% 
  mutate(check=as.character(dense_rank(date))) %>%
  arrange(stream_name) %>% 
  distinct(stream_name, datecount, .keep_all = TRUE) %>%
  spread(check,datecount) %>%
  summarise(`1` = mean(`1`,na.rm=T),
            `2`=mean(`2`,na.rm=T)
            )

phwhall_occobject_1821<-list()

phwhall_occobject_1821$y<-array(dim=c(108,171,2))

phwhall_occobject_1821$y <- array(c(as.matrix(firstcheck_all_1821),as.matrix(secondcheck_all_1821)), dim=c(108,171,2))

phwhall_occobject_1821$det.covs$day<-as.matrix(sitechecks_1821_surveyhistory[,2:3])

dayyear_1821<-cbind(as.numeric(phwh_alltax_suff_env_detecthistory_years_1821_firstcheck$dayofyear),as.numeric(phwh_alltax_suff_env_detecthistory_secondcheck_1821$dayofyear))

phwhall_occobject_1821$det.covs$dayyear<-as.matrix(dayyear_1821)

phwh_subset_suff_env_detecthistory_1821<-phwh_alltax_suff_env_detecthistory_years_1821 %>% 
  dplyr::select(stream_name,reservation,latitude,longitude,river_basin,date,substrate_metric_total,percent_best_types,per_boulder_slabs,per_boulder_256mm,
         per_bedrock,per_cobble,per_gravel,per_sand,per_silt,per_leaf_woody,per_detritus,per_clay_hardpan,per_muck,per_artificial,pool_metric_total,
         pool_depth_.cm.,bankfull_metric_total,average_bankfull_.m.,sinuosity,X._canopy_open,temp,
         dissolved_oxygen_mgl,ph,conductivity,hmfei_narrative) %>%
  group_by(stream_name, reservation,date)

occcov_1821<-phwh_subset_suff_env_detecthistory_1821 %>% group_by(stream_name) %>% 
  summarise_if(is.numeric, mean, na.rm = TRUE)

occcov_1821_2<-occcov_1821[,c(-1,-4)]

cols<-seq(1,25,1)

occcov_1821_2[,cols] = apply(occcov_1821_2[,cols], 2, function(x) as.numeric(as.character(x)))

occcov_1821_3<-occcov_1821_2[ , colSums(is.na(occcov_1821_2)) == 0]

phwhall_occobject_1821$occ.covs<-as.matrix(occcov_1821_3)

library(rgdal)
d <- data.frame(lon=occcov_1821_3$longitude, lat=occcov_1821_3$latitude)
coordinates(d) <- c("lon", "lat")
proj4string(d) <- CRS("+init=epsg:4326") # WGS 84
CRS.new <- CRS("+init=epsg:3734")
d.3734 <- spTransform(d, CRS.new)
plot(d)
projcoords<-data.frame(d.3734)

phwhall_occobject_1821$coords<-as.matrix(projcoords)

rownames(phwhall_occobject_1821$y)<-rownames(firstcheck_all_1821)

```



# RDAs

```{r, warning = FALSE, message= FALSE}
library(vegan)

p_alltax_step_0<-rda(streambioticsum_suff_biotic_bin~1,occcov_3[,-1],scale=T)
p_alltax_step_1<-rda(streambioticsum_suff_biotic_bin~.,occcov_3[,-1],scale=T)

set.seed(12345)
p_alltax_step<- ordistep(p_alltax_step_0, scope = formula(p_alltax_step_1))
p_alltax_step
anova(p_alltax_step)
anova(p_alltax_step,by="term")
library(ggord)

phwh_salamander_suff_sum_bin<-phwh_salamander_suff_sum %>% mutate_if(is.numeric, ~1 * (. >= 1))

phwh_salamander_suff_env_detecthistory<-cbind(phwh_subset_suff_env_detecthistory,phwh_salamander_suff_sum_bin[,-1:-2])


phwh_salamander_suff_env_detecthistory$datecount<-sitechecks$datecount

phwh_salamander_suff_env_detecthistory$dayofyear<-sitechecks$dayofyear



phwh_salamander_suff_env_detecthistory<-phwh_salamander_suff_env_detecthistory %>%
  dplyr::select(!c(latitude,longitude))


phwh_subset_suff_env_detecthistory_first<- phwh_salamander_suff_env_detecthistory %>% group_by(stream_name) %>% 
  arrange(datecount) %>% 
  summarise_all(funs(list(.[1]))) %>% 
  unnest(cols = c(date, reservation, river_basin,hmfei_narrative, substrate_metric_total, percent_best_types, per_boulder_slabs, 
                  per_boulder_256mm, per_bedrock, per_cobble, per_gravel, per_sand, 
                  per_silt, per_leaf_woody, per_detritus, per_clay_hardpan, 
                  per_muck, per_artificial, pool_metric_total, pool_depth_.cm., 
                  bankfull_metric_total, average_bankfull_.m., sinuosity, X._canopy_open, 
                  temp, dissolved_oxygen_mgl, ph, conductivity, mtdusky, nodusky, 
                  twolined, longtail, red, mole, fourtoed, redback, slimy, 
                  datecount, dayofyear))


hmfei_narrative<-phwh_subset_suff_env_detecthistory_first$hmfei_narrative

ggord(p_alltax_step,  xlim=c(-1,1.25),ylim = c(-1.25, 0.75),txt=4,arrow=TRUE,ptslab=TRUE,addsize=3, size =4,grp_in=hmfei_narrative)
```

This sets up a fairly clear contrast between substrate sandiness and most other substrate metrics.

## taxa most differentiated with axes

```{r, warning = FALSE, message= FALSE}
## most differentiated taxa

plot_df <- scores(p_alltax_step, display = "sites") %>% 
  as.data.frame() 
plot_df$stream_name<-streambioticsum_suff$stream_name

plot_df <- plot_df %>%
  full_join(streambioticsum_suff, by = "stream_name")

# envfit() takes the output of metaMDS() and the species matrix you created
fitrdaall <- envfit(p_alltax_step, streambioticsum_suff_biotic_bin, perm = 999) 

# extract p-values for each species
fitrdaall_pvals <- fitrdaall$vectors$pvals %>% 
  as.data.frame() %>% 
  rownames_to_column("species") %>% 
  dplyr::rename("pvals" = ".")

# extract coordinates for species, only keep species with p-val = 0.001
fit_spp <- fitrdaall %>% 
  scores(., display = "vectors") %>% 
  as.data.frame() %>% 
  rownames_to_column("species") %>% 
  full_join(., fitrdaall_pvals, by = "species") %>% 
  filter(pvals == 0.001)

rdaall_plot_new <- ggplot(plot_df, aes(x = RDA1, y = RDA2)) +
  coord_fixed() +
  geom_point(aes(color = hmfei_narrative), size = 3, alpha = 0.8) +
  stat_ellipse(aes(color = hmfei_narrative)) +
  geom_segment(data = fit_spp, aes(x = 0, xend = RDA1, y = 0, yend = RDA2),
               col = "black") +
  geom_text(data = fit_spp, aes(label = species)) 
rdaall_plot_new

fit_spp

```

# cooccurrence networks

## summarizing all taxa

```{r, warning = FALSE, message= FALSE}

phwh_subset_biotic<-phwh %>%
  dplyr::select(stream_name,date,mountain_dusky_larvae,mountain_dusky_juveniles,
                mountain_dusky_adults,northern_dusky_larvae,northern_dusky_juveniles,northern_dusky_adults,two_lined_larvae,two_lined_juveniles,
                two_lined_adults,long_tailed_larvae,long_tailed_juveniles,long_tailed_adults,northern_red_larvae,northern_red_juveniles,northern_red_adults,
                mole_larvae,mole_juveniles,mole_adults,four_toed_larvae,four_toed_juveniles,four_toed_adults,red_backed_juveniles,red_backed_adults,
                slimy_juveniles,slimy_adults,blacknose_dace,bluegill_sunfish,bluntnose_minnow,brindled_madtom,brook_stickleback,brook_trout,
                central_mudminnow,central_stoneroller_minnow,common_shiner,creek_chub,eastern_sand_darter,fantail_darter,fathead_minnow,golden_shiner,
                goldfish,grass_pickerel,greenside_darter,green_sunfish,johnny_darter,largemouth_bass,longear_sunfish,longnose_dace,mottled_sculpin,
                northern_hog_sucker,pumpkinseed_sunfish,rainbow_darter,rainbow_trout,redbelly_dace,redside_dace,river_chub,silverjaw_minnow,smallmouth_bass,
                spotfin_shiner,stonecat_madtom,striped_shiner,trout_perch,warmouth_sunfish,white_sucker,yellow_bullhead_catfish,unidentified_fish_fry,
                sessile_animals,aquatic_worms,sow_bugs,scuds,water_mites,damselfly_nymphs,alderfly_larvae,other_beetles,crayfish,dragonfly_nymphs,
                riffle_beetles,other_flies,midges,snails,clams,fishfly_larvae,water_penny_beetles,cranefly_larvae,ameletidae,arthropleidae,baetidae,
                baetiscidae,caenidae,ephemerellidae,ephemeridae,heptageniidae,isonychiidae,leptohyphidae,leptophlebiidae,polymitarcyidae,potamanthidae,
                pseudironidae,siphlonuridae,capniidae,chloroperlidae,leuctridae,nemouridae,peltoperlidae,perlidae,perlodidae,pteronarcyidae,taeniopterygidae,
                brachycentridae,dipseuodopsidae,glossosomatidae,goeridae,helicopsychidae,hydropsychidae,hydroptilidae,lepidostomatidae,leptoceridae,
                limnephilidae,molannidae,odontoceridae,philopotamidae,phryganeidae,polycentropodidae,psychomiidae,rhyacophilidae,uenoidae) %>% 
  distinct(stream_name, date, .keep_all = TRUE)

phwh_subset_biotic_2<-phwh_subset_biotic %>% replace(is.na(.),0) %>% replace(.=="abundant",30) %>% replace(.=="common",6) %>% replace(.=="rare", 1) %>% replace(.=="very abundant",50) %>%
  distinct(stream_name, date, .keep_all = TRUE) %>% dplyr::select(-stream_name,-date)

phwh_subset_biotic_3<-data.frame(sapply(phwh_subset_biotic_2,as.numeric))

phwh_subset_biotic_3$stream_name<-phwh_subset_biotic$stream_name

phwh_subset_biotic_3$date<-phwh_subset_biotic$date


streambiotic<-phwh_subset_biotic_3
streambioticsum<-streambiotic %>% 
  group_by(stream_name) %>%
  dplyr::select(!date) %>%
  summarise_all(list(sum))

streambioticsum_suff <-streambioticsum[streambioticsum$stream_name %in% greenlightsites$stream_nam , ]

streambioticsum_suff_mtdusky<-streambioticsum_suff$mountain_dusky_larvae+
  streambioticsum_suff$mountain_dusky_juveniles+
  streambioticsum_suff$mountain_dusky_adults

streambioticsum_suff_nodusky<-streambioticsum_suff$northern_dusky_larvae+
  streambioticsum_suff$northern_dusky_juveniles+
  streambioticsum_suff$northern_dusky_adults

streambioticsum_suff_twolined<-streambioticsum_suff$two_lined_larvae+
  streambioticsum_suff$two_lined_juveniles+
  streambioticsum_suff$two_lined_adults

streambioticsum_suff_longtail<-streambioticsum_suff$long_tailed_larvae+
  streambioticsum_suff$long_tailed_juveniles+
  streambioticsum_suff$long_tailed_adults

streambioticsum_suff_red<-streambioticsum_suff$northern_red_larvae+
  streambioticsum_suff$northern_red_juveniles+
  streambioticsum_suff$northern_red_adults

streambioticsum_suff_mole<-streambioticsum_suff$mole_larvae+
  streambioticsum_suff$mole_juveniles+
  streambioticsum_suff$mole_adults

streambioticsum_suff_fourtoed<-streambioticsum_suff$four_toed_larvae+
  streambioticsum_suff$four_toed_juveniles+
  streambioticsum_suff$four_toed_adults

streambioticsum_suff_redback<-streambioticsum_suff$red_backed_juveniles+
  streambioticsum_suff$red_backed_adults

streambioticsum_suff_slimy<-streambioticsum_suff$slimy_juveniles+
  streambioticsum_suff$slimy_adults

streambioticsum_suff_sal<-data.frame(streambioticsum_suff_mtdusky,streambioticsum_suff_nodusky,streambioticsum_suff_twolined,
                                     streambioticsum_suff_longtail,streambioticsum_suff_red,streambioticsum_suff_mole,
                                     streambioticsum_suff_fourtoed,streambioticsum_suff_redback,streambioticsum_suff_slimy)

streambioticsum_suff_fish<-streambioticsum_suff[,27:65]

streambioticsum_suff_macro<-streambioticsum_suff[,67:126]

streambioticsum_suff_biotic<-data.frame(streambioticsum_suff_sal,streambioticsum_suff_fish,streambioticsum_suff_macro)

streambioticsum_suff_biotic_bin<- streambioticsum_suff_biotic %>% mutate_if(is.numeric, ~1 * (. >= 1))

streambioticsum_suff_biotic_bin_t<-t(streambioticsum_suff_biotic_bin)
```

```{r, warning = FALSE, message= FALSE,}

library(netassoc)

m_obs<-matrix(streambioticsum_suff_biotic_bin_t,ncol=137,nrow=108)

m_obs<-m_obs[rowSums(m_obs[])>0,]

m_obs<-m_obs[rowSums(m_obs[])<137,]


# Number of m species
nsp <- nrow(m_obs)
# Number of n sites
nsi <- ncol(m_obs)


m_nul <- floor(matrix(rbernoulli(nsp*nsi,p=0.5),ncol=nsi,nrow=nsp))

n <- make_netassoc_network(m_obs, m_nul,
                           method="partial_correlation",
                           args=list(method="shrinkage"), # for alternative estimators see ?partial_correlation
                           p.method='fdr', 
                           numnulls=100, 
                           plot=TRUE,
                           alpha=0.001)
n$network_all
plot(n$network_all)




# presence/absence across plots

streambioticsum_suff_biotic_bin_t_2<-streambioticsum_suff_biotic_bin_t[rowSums(streambioticsum_suff_biotic_bin_t[])>0,]
streambioticsum_suff_biotic_bin_t_3<-streambioticsum_suff_biotic_bin_t_2[rowSums(streambioticsum_suff_biotic_bin_t_2[])<137,]


library(cooccur)
library(igraph)
co <- print(cooccur(streambioticsum_suff_biotic_bin_t_3, spp_names = TRUE))

co[, "sp1_name"] == rownames(streambioticsum_suff_biotic_bin_t_3)[co$sp1]
co[, "sp2_name"] == rownames(streambioticsum_suff_biotic_bin_t_3)[co$sp2]



nodes <- data.frame(id = 1:nrow(streambioticsum_suff_biotic_bin_t_3),
                    label = rownames(streambioticsum_suff_biotic_bin_t_3),
                    shadow = TRUE) 
edges <- data.frame(from = co$sp1, to = co$sp2,
                    color = ifelse(co$p_lt <= 0.05, "#B0B2C1", "#3C3F51"),
                    dashes = ifelse(co$p_lt <= 0.05, TRUE, FALSE))

cooc<-cooccur(streambioticsum_suff_biotic_bin_t_3, spp_names = TRUE)

visNetwork(nodes = nodes, edges = edges) %>%
  visIgraphLayout(layout = "layout_with_kk")
plot(cooc)

pair.attributes(cooc)
pair.profile(cooc)
prob.table(cooc)


# network stuff

library(asnipe)
library(igraph)
assoc=streambioticsum_suff_biotic_bin_t_3 #transpose the data into group-by-individual
mat=get_network(t(assoc), association_index="SRI") #create adjacency matrix with "simple ratio index"
g.streams=graph_from_adjacency_matrix(mat, "undirected", weighted=T) #make into igraph object
plot(g.streams, edge.width=E(g.streams)$weight*5, vertex.label="")
com=cluster_louvain(g.streams)
com
com$memberships

set.seed(2)
plot(com, g.streams, edge.width=E(g.streams)$weight*0.5)

library(RColorBrewer)
colors=brewer.pal(length(com),'Accent') #make a color palette
V(g.streams)$color=colors[membership(com)] #assign each vertex a color based on the community assignment

set.seed(2)
plot(com,g.streams, edge.width=E(g.streams)$weight*0.1,edge.color="lightgrey",vertex.label.cex=0.8,vertex.size=7,rescale=TRUE)





# number of species
S <- vcount(g.streams)

# number of interactions
L <- ecount(g.streams)

# average number of interactions species
L.S <- L/S

# network connectance
C <- L/S^2

# Calculate connectance
connectance <- ecount(g.streams) / vcount(g.streams)^2 

# Print connectance va
print(paste0('Connectance of stream network =', round(connectance,2)))

links.per.species <- ecount(g.streams) / vcount(g.streams)

## Mean generality of species in the network
Generality <- function(M){
  return(sum(colSums(M))/sum((colSums(M)!=0)));
}
Generality(mat)
  ## Mean vulnerability of the species in the network
  Vulnerability <- function(M){
    return(sum(rowSums(M))/sum((rowSums(M)!=0)));
  }
Vulnerability(mat)
  ## In-degree or number of prey of all species in the network
  InDegree <- function(M){
    return(colSums(M));
  }
InDegree(mat)
  ## Out-degree or number of predators of all species in the network
  OutDegree <- function(M){
    return(rowSums(M));
  }
OutDegree(mat)
```




#occupancy - all taxa

```{r, eval = FALSE, warning = FALSE, message= FALSE}
phwhall_occobject_allsigcor.occ.formula <- ~scale(pool_depth_.cm.) + scale(per_sand) + scale(per_cobble) + scale(substrate_metric_total) +
  scale(pool_metric_total)
phwhall_occobject_season.det.formula <- ~ scale(seasons)

# Number of species
N <- dim(phwhall_occobject$y)[1]
# Distances between sites
dist.phwh <- dist(phwhall_occobject$coords)
# Exponential covariance model
cov.model <- "exponential"
ms.inits <- list(alpha.comm = 0, 
                 beta.comm = 0, 
                 beta = 0, 
                 alpha = 0,
                 tau.sq.beta = 1, 
                 tau.sq.alpha = 1, 
                 z = apply(phwhall_occobject$y, c(1, 2), max, na.rm = TRUE))
ms.priors <- list(beta.comm.normal = list(mean = 0, var = 2.72),
                  alpha.comm.normal = list(mean = 0, var = 2.72), 
                  tau.sq.beta.ig = list(a = 0.1, b = 0.1), 
                  tau.sq.alpha.ig = list(a = 0.1, b = 0.1))


out.alltax.allsigcor.ms <- msPGOcc(occ.formula = phwhall_occobject_allsigcor.occ.formula, 
                                    det.formula = phwhsal_occobject_season.det.formula, 
                                    data = phwhall_occobject, 
                                    inits = ms.inits, 
                                    n.samples = 30000, 
                                    priors = ms.priors, 
                                    n.omp.threads = 4, 
                                    verbose = TRUE, 
                                    n.report = 60000, 
                                    n.burn = 10000,
                                    n.thin = 50, 
                                    n.chains = 3)
summary(out.alltax.allsigcor.ms, level = 'both')
ppc.alltax.allsigcor.ms.out <- ppcOcc(out.alltax.allsigcor.ms, 'chi-squared', group = 1)
summary(ppc.alltax.allsigcor.ms.out, level = 'both')

waicOcc(out.alltax.allsigcor.ms)
```

# improved detection model 

```{r, eval = FALSE, warning = FALSE, message= FALSE}
phwhall_occobject_allsigcor.occ.formula <- ~scale(pool_depth_.cm.) + scale(per_sand) + scale(per_cobble) + scale(substrate_metric_total) +
  scale(pool_metric_total)
phwhall_occobject_dayyearseason.det.formula <- ~scale(seasons)+scale(dayyear)

# Number of species
N <- dim(phwhall_occobject$y)[1]
# Distances between sites
dist.phwh <- dist(phwhall_occobject$coords)
# Exponential covariance model
cov.model <- "exponential"
ms.inits <- list(alpha.comm = 0, 
                 beta.comm = 0, 
                 beta = 0, 
                 alpha = 0,
                 tau.sq.beta = 1, 
                 tau.sq.alpha = 1, 
                 z = apply(phwhall_occobject$y, c(1, 2), max, na.rm = TRUE))
ms.priors <- list(beta.comm.normal = list(mean = 0, var = 2.72),
                  alpha.comm.normal = list(mean = 0, var = 2.72), 
                  tau.sq.beta.ig = list(a = 0.1, b = 0.1), 
                  tau.sq.alpha.ig = list(a = 0.1, b = 0.1))


out.alltax.allsigcor.impdet.ms <- msPGOcc(occ.formula = phwhall_occobject_allsigcor.occ.formula, 
                                    det.formula = phwhall_occobject_dayyearseason.det.formula, 
                                    data = phwhall_occobject, 
                                    inits = ms.inits, 
                                    n.samples = 30000, 
                                    priors = ms.priors, 
                                    n.omp.threads = 4, 
                                    verbose = TRUE, 
                                    n.report = 60000, 
                                    n.burn = 10000,
                                    n.thin = 50, 
                                    n.chains = 3)
summary(out.alltax.allsigcor.impdet.ms, level = 'both')
ppc.alltax.allsigcor.impdet.ms.out <- ppcOcc(out.alltax.allsigcor.impdet.ms, 'chi-squared', group = 1)
summary(ppc.alltax.allsigcor.impdet.ms.out, level = 'both')

waicOcc(out.alltax.allsigcor.impdet.ms)

write_rds(out.alltax.allsigcor.impdet.ms,"alltaxmodel_phwh.rds")

```

# full null

```{r, eval = FALSE, warning = FALSE, message= FALSE}
null <- ~1

# Number of species
N <- dim(phwhall_occobject$y)[1]
# Distances between sites
dist.phwh <- dist(phwhall_occobject$coords)
# Exponential covariance model
cov.model <- "exponential"
ms.inits <- list(alpha.comm = 0, 
                 beta.comm = 0, 
                 beta = 0, 
                 alpha = 0,
                 tau.sq.beta = 1, 
                 tau.sq.alpha = 1, 
                 z = apply(phwhall_occobject$y, c(1, 2), max, na.rm = TRUE))
ms.priors <- list(beta.comm.normal = list(mean = 0, var = 2.72),
                  alpha.comm.normal = list(mean = 0, var = 2.72), 
                  tau.sq.beta.ig = list(a = 0.1, b = 0.1), 
                  tau.sq.alpha.ig = list(a = 0.1, b = 0.1))


out.null <- msPGOcc(occ.formula = null, 
                                    det.formula = null, 
                                    data = phwhall_occobject, 
                                    inits = ms.inits, 
                                    n.samples = 30000, 
                                    priors = ms.priors, 
                                    n.omp.threads = 4, 
                                    verbose = TRUE, 
                                    n.report = 60000, 
                                    n.burn = 10000,
                                    n.thin = 50, 
                                    n.chains = 3)
summary(out.null, level = 'both')
ppc.out.null <- ppcOcc(out.null, 'chi-squared', group = 1)
summary(ppc.out.null, level = 'both')

waicOcc(out.null)
```

Will want to exclude species that did not fit as well. 

```{r, warning = FALSE, message= FALSE}
out.alltax.allsigcor.impdet.ms<-readRDS("alltaxmodel_phwh.rds")

summary(out.alltax.allsigcor.impdet.ms,max.print=100000)
```

# Joint species distribution models

```{r, eval = FALSE, warning = FALSE, message= FALSE}
phwhall_occobject_2<-phwhall_occobject

phwhall_occobject_2$y<-phwhall_occobject$y[(apply(phwhall_occobject$y, 1, mean, na.rm = TRUE) > 0),,]
dim(phwhall_occobject_2$y)


sppnames_2<-rownames(phwhall_occobject_2$y)

sort(apply(phwhall_occobject_2$y, 1, mean, na.rm = TRUE))

# louvain's cluster identified 3 clusters - one of (perhaps) more rocky stream communities, one with most fish that are indicators of # stream health, and another of more spring-associated communities. 

#Place a common species first. The first species has all of its factor loadings set to fixed values, and so it can have a large #influence on the resulting interpretation on the factor loadings and latent factors. We have also found that having a very rare #species first can result in very slow mixing and increased sensitivity to initial values of the latent factor loadings matrix.
#For the remaining q1 factors, place species that you believe will show different occurrence patterns than the first species, and #the other species placed before it. Place these remaining q1 species in order of decreasing differences from the initial factor.

#For example, if I had q=3, for the second species in the array, I would place a species that I a priori think is most different from #the first species. For the third species in the array, I would place a species that I think will show different occurrence patterns #than both the first and second species, but its patterns may not be as noticeably different compared to the first and second species.


# Reorder species. 
sp.ordered <- c("streambiotic_suff_twolined","rainbow_darter","odontoceridae",sppnames_2[!(sppnames_2 %in% c("streambiotic_suff_twolined","rainbow_darter","odontoceridae"))])

#
# Create new detection-nondetection data matrix in the new order
y.new <- phwhall_occobject_2$y[sp.ordered, , ]
# Create a new data array
phwhall.ordered <- phwhall_occobject_2
# Change the data to the new ordered data
phwhall.ordered$y <- y.new
str(phwhall.ordered)
# Number of species
N <- nrow(phwhall.ordered$y)
# Initiate all lambda initial values to 0. 
n.factors<-3
lambda.inits <- matrix(0, N, n.factors)
# Set diagonal elements to 1
diag(lambda.inits) <- 1
# Set lower triangular elements to random values from a standard normal distribution
lambda.inits[lower.tri(lambda.inits)] <- rnorm(sum(lower.tri(lambda.inits)))
# Check it out. Note this is also how spOccupancy specifies default
# initial values for lambda.
lambda.inits

inits <- list(alpha.comm = 0,
              beta.comm = 0,
              beta = 0,
              alpha = 0,
              tau.sq.beta = 1,
              tau.sq.alpha = 1,
              lambda = lambda.inits, 
              z = apply(phwhall.ordered$y, c(1, 2), max, na.rm = TRUE))
priors <- list(beta.comm.normal = list(mean = 0, var = 2.72),
               alpha.comm.normal = list(mean = 0, var = 2.72),
               tau.sq.beta.ig = list(a = 0.1, b = 0.1),
               tau.sq.alpha.ig = list(a = 0.1, b = 0.1))
n.samples <- 5000
n.burn <- 1000
n.thin <- 8
n.chains <- 3
phwhall_occobject_allsigcor.occ.formula <- ~scale(pool_depth_.cm.) + scale(per_sand) + scale(per_cobble) + scale(substrate_metric_total) +
  scale(pool_metric_total)
phwhall_occobject_dayyearseason.det.formula <- ~scale(seasons)+scale(dayyear)
out.lfMsPGOcc <- lfMsPGOcc(occ.formula = phwhall_occobject_allsigcor.occ.formula, det.formula = phwhall_occobject_dayyearseason.det.formula, 
                           data = phwhall.ordered, inits = inits, priors = priors, 
                           n.factors = n.factors, n.samples = n.samples, 
                           n.omp.threads = 1, verbose = TRUE, n.report = 1000, 
                           n.burn = n.burn, n.thin = n.thin, n.chains = n.chains)
summary(out.lfMsPGOcc)
plot(out.lfMsPGOcc$beta.comm.samples, density = FALSE)
summary(out.lfMsPGOcc$lambda.samples)

ppc.out <- ppcOcc(out.lfMsPGOcc, fit.stat = 'freeman-tukey', group = 1)

options(max.print=1000000)
sink("out_lambdas.txt")
print(summary(out.lfMsPGOcc$lambda.samples,quantiles=c(0.05,0.5,0.95)))
sink() 


write.csv(summary(out.lfMsPGOcc$lambda.samples),"out.csv")



```



# now spatial model

```{r, eval = FALSE, warning = FALSE, message= FALSE}
# Pair-wise distance between all sites
dist.phwh <- dist(phwhall_occobject_2$coords)
# Exponential correlation model
cov.model <- "exponential"
# Specify all other initial values identical to lfMsPGOcc() from before
# Number of species
N <- nrow(phwhall_occobject_2$y)
# Initiate all lambda initial values to 0. 
lambda.inits <- matrix(0, N, n.factors)
# Set diagonal elements to 1
diag(lambda.inits) <- 1
# Set lower triangular elements to random values from a standard normal dist
lambda.inits[lower.tri(lambda.inits)] <- rnorm(sum(lower.tri(lambda.inits)))
# Check it out
lambda.inits

inits <- list(alpha.comm = 0,
              beta.comm = 0,
              beta = 0,
              alpha = 0,
              tau.sq.beta = 1,
              tau.sq.alpha = 1,
              lambda = lambda.inits, 
              phi = 3 / mean(dist.phwh),
              z = apply(phwhall.ordered$y, c(1, 2), max, na.rm = TRUE))
batch.length <- 25
n.batch <- 200
n.burn <- 3000
n.thin <- 2
n.chains <- 3
min.dist <- min(dist.phwh)
max.dist <- max(dist.phwh)
priors <- list(beta.comm.normal = list(mean = 0, var = 2.72),
               alpha.comm.normal = list(mean = 0, var = 2.72),
               tau.sq.beta.ig = list(a = 0.1, b = 0.1),
               tau.sq.alpha.ig = list(a = 0.1, b = 0.1), 
               phi.unif = list(3 / max.dist, 3 / min.dist))
tuning <- list(phi = 1)
n.omp.threads <- 1
verbose <- TRUE
n.report <- 50 # Report progress at every 50th batch.

out.sfMsPGOcc <- sfMsPGOcc(occ.formula = phwhall_occobject_allsigcor.occ.formula, det.formula = phwhall_occobject_dayyearseason.det.formula, 
                           data = phwhall.ordered, 
                           inits = inits, 
                           n.batch = n.batch, 
                           batch.length = batch.length, 
                           accept.rate = 0.43, 
                           priors = priors, 
                           n.factors = n.factors,
                           cov.model = cov.model, 
                           tuning = tuning, 
                           n.omp.threads = n.omp.threads, 
                           verbose = TRUE, 
                           NNGP = TRUE, 
                           n.neighbors = 5, 
                           n.report = n.report, 
                           n.burn = n.burn, 
                           n.thin = n.thin, 
                           n.chains = n.chains)
summary(out.sfMsPGOcc)
plot(out.sfMsPGOcc$beta.comm.samples, density = FALSE)
summary(out.sfMsPGOcc$lambda.samples)

ppc.out.sf <- ppcOcc(out.sfMsPGOcc, fit.stat = 'freeman-tukey', group = 1)
waicOcc(out.sfMsPGOcc)
waicOcc(out.lfMsPGOcc)


lambda.inits.1 <- matrix(0, N, 1)
# Set diagonal elements to 1
diag(lambda.inits.1) <- 1
# Set lower triangular elements to random values from a standard normal dist
lambda.inits.1[lower.tri(lambda.inits.1)] <- rnorm(sum(lower.tri(lambda.inits.1)))
# Check it out
lambda.inits.1
inits.1 <- list(alpha.comm = 0,
              beta.comm = 0,
              beta = 0,
              alpha = 0,
              tau.sq.beta = 1,
              tau.sq.alpha = 1,
              lambda = lambda.inits.1, 
              phi = 3 / mean(dist.phwh),
              z = apply(phwhall.ordered$y, c(1, 2), max, na.rm = TRUE))

out.sfMsPGOcc.fac1 <- sfMsPGOcc(occ.formula = phwhall_occobject_allsigcor.occ.formula, det.formula = phwhall_occobject_dayyearseason.det.formula, 
                           data = phwhall.ordered, 
                           inits = inits.1, 
                           n.batch = n.batch, 
                           batch.length = batch.length, 
                           accept.rate = 0.43, 
                           priors = priors, 
                           n.factors = 1,
                           cov.model = cov.model, 
                           tuning = tuning, 
                           n.omp.threads = n.omp.threads, 
                           verbose = TRUE, 
                           NNGP = TRUE, 
                           n.neighbors = 5, 
                           n.report = n.report, 
                           n.burn = n.burn, 
                           n.thin = n.thin, 
                           n.chains = n.chains)

summary(out.sfMsPGOcc.fac1$lambda.samples)



```

This revised model (seems) to point towards sow bugs as *particularly* differentiated from good stream salamander habitat. Let's explore that further - using sow bugs as the second factor.

```{r, eval = FALSE, warning = FALSE, message= FALSE}

# Reorder species. 
sp.ordered.2 <- c("streambiotic_suff_twolined","sow_bugs",sppnames_2[!(sppnames_2 %in% c("streambiotic_suff_twolined","sow_bugs"))])

#
# Create new detection-nondetection data matrix in the new order
y.new.2 <- phwhall_occobject_2$y[sp.ordered.2, , ]
# Create a new data array
phwhall.ordered.2 <- phwhall_occobject_2
# Change the data to the new ordered data
phwhall.ordered.2$y <- y.new.2
str(phwhall.ordered.2)
# Number of species
N <- nrow(phwhall.ordered.2$y)
# Initiate all lambda initial values to 0. 
n.factors<-2
lambda.inits <- matrix(0, N, n.factors)
# Set diagonal elements to 1
diag(lambda.inits) <- 1
# Set lower triangular elements to random values from a standard normal distribution
lambda.inits[lower.tri(lambda.inits)] <- rnorm(sum(lower.tri(lambda.inits)))
# Check it out. Note this is also how spOccupancy specifies default
# initial values for lambda.
lambda.inits

n.samples <- 5000
n.burn <- 1000
n.thin <- 8
n.chains <- 3
phwhall_occobject_allsigcor.occ.formula <- ~scale(pool_depth_.cm.) + scale(per_sand) + scale(per_cobble) + scale(substrate_metric_total) +
  scale(pool_metric_total)
phwhall_occobject_dayyearseason.det.formula <- ~scale(seasons)+scale(dayyear)
inits.2 <- list(alpha.comm = 0,
              beta.comm = 0,
              beta = 0,
              alpha = 0,
              tau.sq.beta = 1,
              tau.sq.alpha = 1,
              lambda = lambda.inits, 
              phi = 3 / mean(dist.phwh),
              z = apply(phwhall.ordered.2$y, c(1, 2), max, na.rm = TRUE))

out.sfMsPGOcc.fac2 <- sfMsPGOcc(occ.formula = phwhall_occobject_allsigcor.occ.formula, det.formula = phwhall_occobject_dayyearseason.det.formula, 
                           data = phwhall.ordered.2, 
                           inits = inits.2, 
                           n.batch = n.batch, 
                           batch.length = batch.length, 
                           accept.rate = 0.43, 
                           priors = priors, 
                           n.factors = n.factors,
                           cov.model = cov.model, 
                           tuning = tuning, 
                           n.omp.threads = n.omp.threads, 
                           verbose = TRUE, 
                           NNGP = TRUE, 
                           n.neighbors = 5, 
                           n.report = n.report, 
                           n.burn = n.burn, 
                           n.thin = n.thin, 
                           n.chains = n.chains)

summary(out.sfMsPGOcc.fac2$lambda.samples)

waicOcc(out.sfMsPGOcc.fac1)

waicOcc(out.sfMsPGOcc.fac2)


```

Nope - q = 1 is still best. 

out.sfMsPGOcc.fac1 should likely be the model to use for now.

reasonable fit for this model already! let's try to predict out to all streams.

```{r, eval = FALSE, warning = FALSE, message= FALSE}

phwh_subset_2<-phwh_subset %>% dplyr::select(stream_name,date,latitude,longitude,pool_depth_.cm.,per_sand,per_cobble,substrate_metric_total,pool_metric_total) %>% distinct(stream_name, date, .keep_all = TRUE)
  
phwh_subset_3<-na.omit(phwh_subset_2)


pool_depth_.cm. = ((phwh_subset_3$pool_depth_.cm. - mean(phwhall.ordered$occ.covs[,16]))/sd(phwhall.ordered$occ.covs[,16]))
per_sand = ((phwh_subset_3$per_sand - mean(phwhall.ordered$occ.covs[,8]))/sd(phwhall.ordered$occ.covs[,8]))
per_cobble = ((phwh_subset_3$per_cobble - mean(phwhall.ordered$occ.covs[, 6])) / sd(phwhall.ordered$occ.covs[, 6]))
substrate_metric_total = ((phwh_subset_3$substrate_metric_total - mean(phwhall.ordered$occ.covs[, 1])) / sd(phwhall.ordered$occ.covs[, 1]))
pool_metric_total = ((as.numeric(phwh_subset_3$pool_metric_total) - mean(phwhall.ordered$occ.covs[, 15])) / sd(phwhall.ordered$occ.covs[, 15]))

predict_data<-data.frame(int = 1, pool_depth_.cm. = pool_depth_.cm., per_sand = per_sand, per_cobble = per_cobble, substrate_metric_total = substrate_metric_total, pool_metric_total = pool_metric_total)

coords<-as.matrix(phwh_subset_3[,c("longitude","latitude")])

fullphwhpred<-predict(out.sfMsPGOcc.fac1,as.matrix(predict_data),coords,type="occupancy")

str(fullphwhpred)
# Species richness samples
rich.pred <- apply(fullphwhpred$z.0.samples, c(1, 3), sum)
rich.real<-apply(na.omit(phwhall_occobject$y),c(2,3),function(x) sum(x,na.rm=T))
latersurveys<-rich.real[,3:8]
latersurveys[latersurveys==0]<-NA
rich.real[,3:8]<-latersurveys
plot.dat <- data.frame(stream_name=phwh_subset_3$stream_name,
                       date=phwh_subset_3$date,
                       x = phwh_subset_3$longitude, 
                       y = phwh_subset_3$latitude, 
                       rich.mean = apply(rich.pred, 2, mean), 
                       rich.sd = apply(rich.pred, 2, sd))

plot.dat.real <- data.frame(stream_name=colnames(seventhcheck_all_2),
                       x = phwhall_occobject$coords[,1], 
                       y = phwhall_occobject$coords[,2], 
                       rich.mean = apply(rich.real, 1, function(x) mean(x,na.rm=T)), 
                       rich.sd = apply(rich.real, 1, function(x) sd(x,na.rm=T)))
# Plot species richness across PHWH

RMSE <- function(observed, predicted) {
  sqrt(mean((predicted - observed)^2, na.rm=TRUE))
}
f1 <- function(x, test, train) {
  nmx <- x[1]
  idp <- x[2]
  if (nmx < 1) return(Inf)
  if (idp < .001) return(Inf)
  m <- gstat(formula=rich.mean~1, locations=train, nmax=nmx, set=list(idp=idp))
  p <- predict(m, newdata=test, debug.level=0)$var1.pred
  RMSE(test$rich.mean, p)
}

set.seed(12345)


respoly<-readOGR("G:/NaturalResources/Byer/DataAnalysis/bbs_cmp_analyses/bbsapp/CMPboundaries.shp")

respoly <- spTransform(respoly, CRS("+init=epsg:4326"))

plot.dat$date<-as.Date(plot.dat$date , format = "%Y/%m/%d")

plot.dat$year<-year(plot.dat$date)

sort(unique(plot.dat$year))

plot.dat_20032010<-plot.dat[plot.dat$year %in% c(2003,2004,2007,2008,20009,2010),]

plot.dat_20172022<-plot.dat[plot.dat$year %in% c(2017,2018,2019,2020,2021,2022),]
  
  dsp <- SpatialPoints(plot.dat_20172022[,3:4])
  dsp <- SpatialPointsDataFrame(dsp, plot.dat_20172022)
  crs(dsp)<-CRS("+init=epsg:4326")
  j <- sample(nrow(dsp), 0.2 * nrow(dsp))
  tst <- dsp[j,]
  trn <- dsp[-j,]
  opt <- optim(c(8, .5), f1, test=tst, train=trn)
  
  # define groups for mapping
  cuts <- c(14,21,28,35)
  # set up a palette of interpolated colors
  blues <- colorRampPalette(c('yellow', 'orange', 'blue', 'dark blue'))
  spplot(dsp, 'rich.mean', cuts=cuts, col.regions=blues(5), pch=20, cex=2)
  v <- voronoi(dsp)
  
  ch <- chull(plot.dat_20172022[,3:4])
  coords <- plot.dat_20172022[c(ch, ch[1]), 3:4]  # closed polygon
  
  plot(plot.dat_20172022[,3:4], pch=19)
  lines(coords, col="red")
  
 
  vca <- intersect(v, respoly)
  spplot(vca, 'rich.mean', col.regions=rev(get_col_regions()))
  r <- raster(respoly, res=0.0001)
  crs(r)<-CRS("+init=epsg:4326")
  crs(vca)<-CRS("+init=epsg:4326")
  vr <- rasterize(vca, r, 'rich.mean')
  crs(vr)<-CRS("+init=epsg:4326")
  plot(vr)
  points(dsp)
  
  writeRaster(vr,filename = paste0(getwd(),"richness_20172022.asc"),overwrite=TRUE)
  
  #actual richness
  
  dsp <- SpatialPoints(plot.dat.real[,2:3])
  dsp <- SpatialPointsDataFrame(dsp, plot.dat.real)
  crs(dsp)<-CRS("+init=epsg:3734")
  j <- sample(nrow(dsp), 0.2 * nrow(dsp))
  tst <- dsp[j,]
  trn <- dsp[-j,]
  opt <- optim(c(8, .5), f1, test=tst, train=trn)
 
  respoly.3734<-spTransform(respoly,CRS("+init=epsg:3734"))
  r <- raster(respoly.3734, res=300)
  
  fitmax <- gstat::gstat(formula = rich.mean ~ 1, data = dsp, nmax = opt$par[1], set = list(idp = opt$par[2]))
  maxint <- raster::interpolate(r, model=fitmax)
  crs(maxint)<-CRS("+init=epsg:3734")
  plot(maxint, col=rev(heat.colors(255)))
  r2 <- crop(maxint, extent(respoly.3734))
  r3 <- mask(r2, respoly.3734)
    plot(r3, col=rev(heat.colors(255)))

  
  writeRaster(r3,filename = paste0(getwd(),"richness_observed.asc"),overwrite=TRUE)

```

As a brief note - all of the above assumes closure (no immigration/emigration across surveys), which is likely violated. There are two suites of approaches that could be used at this juncture - Hmsc, assuming perfect detection, and Fidino et al. (2019)'s dcom. First, I will try just fitting the spatial factor occupancy model to 2008 data.

```{r, eval = FALSE, warning = FALSE, message= FALSE}
phwhall_occobject_2008_2<-phwhall_occobject_2008

phwhall_occobject_2008_2$y<-phwhall_occobject_2008$y[(apply(phwhall_occobject_2008$y, 1, mean, na.rm = TRUE) > 0),,]
dim(phwhall_occobject_2008_2$y)


sppnames_2008_2<-rownames(phwhall_occobject_2008_2$y)

sort(apply(phwhall_occobject_2008_2$y, 1, mean, na.rm = TRUE))

# louvain's cluster identified 3 clusters - one of (perhaps) more rocky stream communities, one with most fish that are indicators of # stream health, and another of more spring-associated communities. 

#Place a common species first. The first species has all of its factor loadings set to fixed values, and so it can have a large #influence on the resulting interpretation on the factor loadings and latent factors. We have also found that having a very rare #species first can result in very slow mixing and increased sensitivity to initial values of the latent factor loadings matrix.
#For the remaining q1 factors, place species that you believe will show different occurrence patterns than the first species, and #the other species placed before it. Place these remaining q1 species in order of decreasing differences from the initial factor.

#For example, if I had q=3, for the second species in the array, I would place a species that I a priori think is most different from #the first species. For the third species in the array, I would place a species that I think will show different occurrence patterns #than both the first and second species, but its patterns may not be as noticeably different compared to the first and second species.


# Reorder species. 
sp.2008.ordered <- c("midges","streambiotic_suff_mole","johnny_darter",sppnames_2008_2[!(sppnames_2008_2 %in% c("midges","streambiotic_suff_mole","johnny_darter"))])

# Create new detection-nondetection data matrix in the new order
y.2008.new <- phwhall_occobject_2008_2$y[sp.2008.ordered, , ]
# Create a new data array
phwhall.2008.ordered <- phwhall_occobject_2008_2
# Change the data to the new ordered data
phwhall.2008.ordered$y <- y.2008.new
str(phwhall.2008.ordered)

# Pair-wise distance between all sites
dist.phwh <- dist(phwhall.2008.ordered$coords)
# Exponential correlation model
cov.model <- "exponential"
# Specify all other initial values identical to lfMsPGOcc() from before
# Number of species
N <- nrow(phwhall.2008.ordered$y)
# Initiate all lambda initial values to 0. 
lambda.inits <- matrix(0, N, n.factors)
# Set diagonal elements to 1
diag(lambda.inits) <- 1
# Set lower triangular elements to random values from a standard normal dist
lambda.inits[lower.tri(lambda.inits)] <- rnorm(sum(lower.tri(lambda.inits)))
# Check it out
lambda.inits

inits <- list(alpha.comm = 0,
              beta.comm = 0,
              beta = 0,
              alpha = 0,
              tau.sq.beta = 1,
              tau.sq.alpha = 1,
              lambda = lambda.inits, 
              phi = 3 / mean(dist.phwh),
              z = apply(phwhall.2008.ordered$y, c(1, 2), max, na.rm = TRUE))
batch.length <- 50
n.batch <- 500
n.burn <- 5000
n.thin <- 2
n.chains <- 3
min.dist <- min(dist.phwh)
max.dist <- max(dist.phwh)
priors <- list(beta.comm.normal = list(mean = 0, var = 2.72),
               alpha.comm.normal = list(mean = 0, var = 2.72),
               tau.sq.beta.ig = list(a = 0.1, b = 0.1),
               tau.sq.alpha.ig = list(a = 0.1, b = 0.1), 
               phi.unif = list(3 / max.dist, 3 / min.dist))
tuning <- list(phi = 1)
n.omp.threads <- 1
verbose <- TRUE
n.report <- 50 # Report progress at every 50th batch.

phwhall_occobject_2008_allsigcor.occ.formula <- ~scale(substrate_metric_total) +
  scale(pool_metric_total)
phwhall_occobject_2008_dayyearseason.det.formula <- ~scale(dayyear)
out.sfMsPGOcc.2008 <- sfMsPGOcc(occ.formula = phwhall_occobject_2008_allsigcor.occ.formula, det.formula = phwhall_occobject_2008_dayyearseason.det.formula, 
                           data = phwhall.2008.ordered, 
                           inits = inits, 
                           n.batch = n.batch, 
                           batch.length = batch.length, 
                           accept.rate = 0.43, 
                           priors = priors, 
                           n.factors = n.factors,
                           cov.model = cov.model, 
                           tuning = tuning, 
                           n.omp.threads = n.omp.threads, 
                           verbose = TRUE, 
                           NNGP = TRUE, 
                           n.neighbors = 5, 
                           n.report = n.report, 
                           n.burn = n.burn, 
                           n.thin = n.thin, 
                           n.chains = n.chains)
summary(out.sfMsPGOcc.2008)
plot(out.sfMsPGOcc.2008$beta.comm.samples, density = FALSE)
summary(out.sfMsPGOcc.2008$lambda.samples)

ppc.out.2008 <- ppcOcc(out.sfMsPGOcc.2008, fit.stat = 'freeman-tukey', group = 1)
summary(ppc.out.2008)

```

# improved detection model - for 2008 alone

```{r, eval = FALSE, warning = FALSE, message= FALSE}
phwhall_occobject_allsigcor.occ.formula <- ~scale(pool_depth_.cm.) + scale(per_sand) + scale(per_cobble) + scale(substrate_metric_total) +
  scale(pool_metric_total)
phwhall_occobject_dayyear.det.formula <- ~scale(dayyear)

# Number of species
N <- dim(phwhall.2008.ordered$y)[1]
# Distances between sites
dist.phwh <- dist(phwhall.2008.ordered$coords)
# Exponential covariance model
cov.model <- "exponential"
ms.inits <- list(alpha.comm = 0, 
                 beta.comm = 0, 
                 beta = 0, 
                 alpha = 0,
                 tau.sq.beta = 1, 
                 tau.sq.alpha = 1, 
                 z = apply(phwhall.2008.ordered$y, c(1, 2), max, na.rm = TRUE))
ms.priors <- list(beta.comm.normal = list(mean = 0, var = 2.72),
                  alpha.comm.normal = list(mean = 0, var = 2.72), 
                  tau.sq.beta.ig = list(a = 0.1, b = 0.1), 
                  tau.sq.alpha.ig = list(a = 0.1, b = 0.1))


out.alltax.allsigcor.impdet.ms.2008 <- msPGOcc(occ.formula = phwhall_occobject_allsigcor.occ.formula, 
                                    det.formula = phwhall_occobject_dayyear.det.formula, 
                                    data = phwhall.2008.ordered, 
                                    inits = ms.inits, 
                                    n.samples = 30000, 
                                    priors = ms.priors, 
                                    n.omp.threads = 4, 
                                    verbose = TRUE, 
                                    n.report = 10000, 
                                    n.burn = 5000,
                                    n.thin = 50, 
                                    n.chains = 3)
summary(out.alltax.allsigcor.impdet.ms.2008, level = 'both')
ppc.alltax.allsigcor.impdet.ms.2008.out <- ppcOcc(out.alltax.allsigcor.impdet.ms.2008, 'chi-squared', group = 1)
summary(ppc.alltax.allsigcor.impdet.ms.2008.out, level = 'both')

waicOcc(out.alltax.allsigcor.impdet.ms.2008)

write_rds(out.alltax.allsigcor.impdet.ms,"alltaxmodel_phwh.rds")

```

# occupancy model for 0710 alone

```{r, eval = FALSE, warning = FALSE, message= FALSE}

phwhall_occobject_0710_2<-phwhall_occobject_0710

phwhall_occobject_0710_2$y<-phwhall_occobject_0710$y[(apply(phwhall_occobject_0710$y, 1, mean, na.rm = TRUE) > 0),,]
dim(phwhall_occobject_0710_2$y)


sppnames_0710_2<-rownames(phwhall_occobject_0710_2$y)

sort(apply(phwhall_occobject_0710_2$y, 1, mean, na.rm = TRUE))

# louvain's cluster identified 3 clusters - one of (perhaps) more rocky stream communities, one with most fish that are indicators of # stream health, and another of more spring-associated communities. 

#Place a common species first. The first species has all of its factor loadings set to fixed values, and so it can have a large #influence on the resulting interpretation on the factor loadings and latent factors. We have also found that having a very rare #species first can result in very slow mixing and increased sensitivity to initial values of the latent factor loadings matrix.
#For the remaining q1 factors, place species that you believe will show different occurrence patterns than the first species, and #the other species placed before it. Place these remaining q1 species in order of decreasing differences from the initial factor.

#For example, if I had q=3, for the second species in the array, I would place a species that I a priori think is most different from #the first species. For the third species in the array, I would place a species that I think will show different occurrence patterns #than both the first and second species, but its patterns may not be as noticeably different compared to the first and second species.


# Reorder species. 
sp.0710.ordered <- c("midges","streambiotic_suff_mole","johnny_darter",sppnames_0710_2[!(sppnames_0710_2 %in% c("midges","streambiotic_suff_mole","johnny_darter"))])

# Create new detection-nondetection data matrix in the new order
y.0710.new <- phwhall_occobject_0710_2$y[sp.0710.ordered, , ]
# Create a new data array
phwhall.0710.ordered <- phwhall_occobject_0710_2
# Change the data to the new ordered data
phwhall.0710.ordered$y <- y.0710.new
str(phwhall.0710.ordered)

phwhall_occobject_allsigcor.occ.formula <- ~scale(pool_depth_.cm.) + scale(per_sand) + scale(per_cobble) + scale(substrate_metric_total) +
  scale(pool_metric_total)
phwhall_occobject_dayyear.det.formula <- ~scale(dayyear)

# Number of species
N <- dim(phwhall.0710.ordered$y)[1]
# Distances between sites
dist.phwh <- dist(phwhall.0710.ordered$coords)
# Exponential covariance model
cov.model <- "exponential"
ms.inits <- list(alpha.comm = 0, 
                 beta.comm = 0, 
                 beta = 0, 
                 alpha = 0,
                 tau.sq.beta = 1, 
                 tau.sq.alpha = 1, 
                 z = apply(phwhall.0710.ordered$y, c(1, 2), max, na.rm = TRUE))
ms.priors <- list(beta.comm.normal = list(mean = 0, var = 2.72),
                  alpha.comm.normal = list(mean = 0, var = 2.72), 
                  tau.sq.beta.ig = list(a = 0.1, b = 0.1), 
                  tau.sq.alpha.ig = list(a = 0.1, b = 0.1))


out.alltax.allsigcor.impdet.ms.0710 <- msPGOcc(occ.formula = phwhall_occobject_allsigcor.occ.formula, 
                                    det.formula = phwhall_occobject_dayyear.det.formula, 
                                    data = phwhall.0710.ordered, 
                                    inits = ms.inits, 
                                    n.samples = 30000, 
                                    priors = ms.priors, 
                                    n.omp.threads = 4, 
                                    verbose = TRUE, 
                                    n.report = 10000, 
                                    n.burn = 5000,
                                    n.thin = 50, 
                                    n.chains = 3)
summary(out.alltax.allsigcor.impdet.ms.0710, level = 'both')
ppc.alltax.allsigcor.impdet.ms.0710.out <- ppcOcc(out.alltax.allsigcor.impdet.ms.0710, 'chi-squared', group = 1)
summary(ppc.alltax.allsigcor.impdet.ms.0710.out, level = 'both')

waicOcc(out.alltax.allsigcor.impdet.ms.0710)

write_rds(out.alltax.allsigcor.impdet.ms.0710,"alltaxmodel_phwh_0710.rds")

```

## prediction across census streams - taxa richness?

```{r, eval = FALSE, warning = FALSE, message= FALSE}

rich.pred<-apply(out.alltax.allsigcor.impdet.ms.0710$z.samples, c(1,3), sum)


 plot.dat.rich <- data.frame(x = pcap_full_land$xcoord, 
                            y = pcap_full_land$ycoord, 
                            mean.rich = apply(rich.pred, 2, mean),
                            sd.rich = apply(rich.pred, 2, sd),
                            mean.rich.R = apply(rich.pred.R,2,mean),
                            sd.rich.R = apply(rich.pred.R,2,sd),
                            mean.rich.S = apply(rich.pred.S,2,mean),
                            sd.rich.S = apply(rich.pred.S,2,sd),
                            mean.rich.T = apply(rich.pred.T,2,mean),
                            sd.rich.T = apply(rich.pred.T,2,sd),
                            stringsAsFactors = FALSE)
 
 write.csv(plot.dat.rich,"plotdatarichness.csv")
  
  dsp <- SpatialPoints(plot.dat.rich[,1:2])
  dsp <- SpatialPointsDataFrame(dsp, plot.dat.rich)
  crs(dsp)<-"+proj=lcc +lat_0=39.6666666666667 +lon_0=-82.5 +lat_1=41.7 +lat_2=40.4333333333333 +x_0=600000 +y_0=0 +datum=NAD83 +units=us-ft +no_defs"
  j <- sample(nrow(dsp), 0.2 * nrow(dsp))
  tst <- dsp[j,]
  trn <- dsp[-j,]
RMSE <- function(observed, predicted) {
  sqrt(mean((predicted - observed)^2, na.rm=TRUE))
}
f1 <- function(x, test, train) {
  nmx <- x[1]
  idp <- x[2]
  if (nmx < 1) return(Inf)
  if (idp < .001) return(Inf)
  m <- gstat(formula=mean.rich.S~1, locations=train, nmax=nmx, set=list(idp=idp))
  p <- predict(m, newdata=test, debug.level=0)$var1.pred
  RMSE(test$mean.rich.S, p)
}
p <- data.frame(geom(dsp)[, c("x", "y")], as.data.frame(dsp))
r <- raster(respoly, res=100)
opt <- optim(c(8, 0.5), f1, test=tst, train=trn)
m <- gstat(formula=mean.rich.S~1, locations=~x+y, data=p, nmax=opt$par[1], set=list(idp=opt$par[2]))
idw <- interpolate(r, m, debug.level=0)


idw <- mask(idw, respoly)
plot(idw, 1)
```

## occupancy predictions

```{r, eval = FALSE, warning = FALSE, message= FALSE}
respoly<-readOGR("G:/NaturalResources/Byer/DataAnalysis/bbs_cmp_analyses/bbsapp/CMPboundaries.shp")

RMSE <- function(observed, predicted) {
  sqrt(mean((predicted - observed)^2, na.rm=TRUE))
}
f1 <- function(x, test, train) {
  nmx <- x[1]
  idp <- x[2]
  if (nmx < 1) return(Inf)
  if (idp < .001) return(Inf)
  m <- gstat(formula=mean.psi~1, locations=train, nmax=nmx, set=list(idp=idp))
  p <- predict(m, newdata=test, debug.level=0)$var1.pred
  RMSE(test$mean.psi, p)
}

idwrasters_occupancy_census<-list()

sppnames<-rownames(phwhall.0710.ordered$y)
saveRDS(sppnames,"sppnames_census.RData")
for(i in 1:length(sppnames)){
  plot.dat.ms <- data.frame(x = phwhall.0710.ordered$coords[,1], 
                            y = phwhall.0710.ordered$coords[,2], 
                            mean.psi = apply(out.alltax.allsigcor.impdet.ms.0710$psi.samples[,i,], 2, mean))
  
  dsp <- SpatialPoints(plot.dat.ms[,1:2])
  dsp <- SpatialPointsDataFrame(dsp, plot.dat.ms)
  crs(dsp)<-CRS("+init=epsg:3734")
  j <- sample(nrow(dsp), 0.2 * nrow(dsp))
  tst <- dsp[j,]
  trn <- dsp[-j,]
opt <- optim(c(8, .5), f1, test=tst, train=trn)
  # define groups for mapping
  cuts <- c(0,0.25,0.75,1.0)
  # set up a palette of interpolated colors
  blues <- colorRampPalette(c('yellow', 'orange', 'blue', 'dark blue'))
  spplot(dsp, 'mean.psi', cuts=cuts, col.regions=blues(5), pch=20, cex=2)
  v <- voronoi(dsp)
   
  ch <- chull(plot.dat.ms[,1:2])
  coords <- plot.dat.ms[c(ch, ch[1]), 1:2]  # closed polygon
  
  plot(plot.dat.ms[,1:2], pch=19)
  lines(coords, col="red")
  
  r <- raster(respoly, res=500)
  
  fitmax <- gstat::gstat(formula = mean.psi ~ 1, data = dsp, nmax = opt$par[1], set = list(idp = opt$par[2]))
  maxint <- raster::interpolate(r, model=fitmax)
  crs(maxint)<-"+proj=lcc +lat_0=39.6666666666667 +lon_0=-82.5 +lat_1=41.7 +lat_2=40.4333333333333 +x_0=600000 +y_0=0 +datum=NAD83 +units=us-ft +no_defs"
  plot(maxint, col=rev(heat.colors(255)))
  plot(vr, add=TRUE)
  r2 <- crop(maxint, extent(respoly))
  r3 <- mask(r2, respoly)
  plot(r3)
  idwrasters_occupancy_census[[i]]<-r3
  writeRaster(r3,filename = paste0("G:/NaturalResources/Byer/DataAnalysis/PHWH/rasterpredictions/02082023_predictions/census/",sppnames[i],"_census_meanpsi_02082023.asc"),overwrite=TRUE)
  
}

saveRDS(idwrasters_occupancy_census,"idwrasters_occupancy_census.rData")

options(max.print=1000000)
sink("out.alltax.allsigcor.impdet.ms.0710.txt")
print(summary(out.alltax.allsigcor.impdet.ms.0710,level="species",quantiles=c(0.05,0.5,0.95)))
sink() 

```


# occupancy model for 1316 alone

```{r, eval = FALSE, warning = FALSE, message= FALSE}

phwhall_occobject_1316_2<-phwhall_occobject_1316

phwhall_occobject_1316_2$y<-phwhall_occobject_1316$y[(apply(phwhall_occobject_1316$y, 1, mean, na.rm = TRUE) > 0),,]
dim(phwhall_occobject_1316_2$y)


sppnames_1316_2<-rownames(phwhall_occobject_1316_2$y)

sort(apply(phwhall_occobject_1316_2$y, 1, mean, na.rm = TRUE))

# louvain's cluster identified 3 clusters - one of (perhaps) more rocky stream communities, one with most fish that are indicators of # stream health, and another of more spring-associated communities. 

#Place a common species first. The first species has all of its factor loadings set to fixed values, and so it can have a large #influence on the resulting interpretation on the factor loadings and latent factors. We have also found that having a very rare #species first can result in very slow mixing and increased sensitivity to initial values of the latent factor loadings matrix.
#For the remaining q1 factors, place species that you believe will show different occurrence patterns than the first species, and #the other species placed before it. Place these remaining q1 species in order of decreasing differences from the initial factor.

#For example, if I had q=3, for the second species in the array, I would place a species that I a priori think is most different from #the first species. For the third species in the array, I would place a species that I think will show different occurrence patterns #than both the first and second species, but its patterns may not be as noticeably different compared to the first and second species.


# Reorder species. 
sp.1316.ordered <- c("other_flies","central_stoneroller_minnow","johnny_darter",sppnames_1316_2[!(sppnames_1316_2 %in% c("other_flies","central_stoneroller_minnow","johnny_darter"))])

# Create new detection-nondetection data matrix in the new order
y.1316.new <- phwhall_occobject_1316_2$y[sp.1316.ordered, , ]
# Create a new data array
phwhall.1316.ordered <- phwhall_occobject_1316_2
# Change the data to the new ordered data
phwhall.1316.ordered$y <- y.1316.new
str(phwhall.1316.ordered)

phwhall_occobject_allsigcor.occ.formula <- ~scale(pool_depth_.cm.) + scale(per_sand) + scale(per_cobble) + scale(substrate_metric_total) +
  scale(pool_metric_total)
phwhall_occobject_dayyear.det.formula <- ~scale(dayyear)

# Number of species
N <- dim(phwhall.1316.ordered$y)[1]
# Distances between sites
dist.phwh <- dist(phwhall.1316.ordered$coords)
# Exponential covariance model
cov.model <- "exponential"
ms.inits <- list(alpha.comm = 0, 
                 beta.comm = 0, 
                 beta = 0, 
                 alpha = 0,
                 tau.sq.beta = 1, 
                 tau.sq.alpha = 1, 
                 z = apply(phwhall.1316.ordered$y, c(1, 2), max, na.rm = TRUE))
ms.priors <- list(beta.comm.normal = list(mean = 0, var = 2.72),
                  alpha.comm.normal = list(mean = 0, var = 2.72), 
                  tau.sq.beta.ig = list(a = 0.1, b = 0.1), 
                  tau.sq.alpha.ig = list(a = 0.1, b = 0.1))


out.alltax.allsigcor.impdet.ms.1316 <- msPGOcc(occ.formula = phwhall_occobject_allsigcor.occ.formula, 
                                    det.formula = phwhall_occobject_dayyear.det.formula, 
                                    data = phwhall.1316.ordered, 
                                    inits = ms.inits, 
                                    n.samples = 30000, 
                                    priors = ms.priors, 
                                    n.omp.threads = 4, 
                                    verbose = TRUE, 
                                    n.report = 10000, 
                                    n.burn = 5000,
                                    n.thin = 50, 
                                    n.chains = 3)
summary(out.alltax.allsigcor.impdet.ms.1316, level = 'both')
ppc.alltax.allsigcor.impdet.ms.1316.out <- ppcOcc(out.alltax.allsigcor.impdet.ms.1316, 'chi-squared', group = 1)
summary(ppc.alltax.allsigcor.impdet.ms.1316.out, level = 'both')

waicOcc(out.alltax.allsigcor.impdet.ms.1316)

write_rds(out.alltax.allsigcor.impdet.ms.1316,"alltaxmodel_phwh_1316.rds")

```

## occupancy predictions

```{r, eval = FALSE, warning = FALSE, message= FALSE}
respoly<-readOGR("G:/NaturalResources/Byer/DataAnalysis/bbs_cmp_analyses/bbsapp/CMPboundaries.shp")

RMSE <- function(observed, predicted) {
  sqrt(mean((predicted - observed)^2, na.rm=TRUE))
}
f1 <- function(x, test, train) {
  nmx <- x[1]
  idp <- x[2]
  if (nmx < 1) return(Inf)
  if (idp < .001) return(Inf)
  m <- gstat(formula=mean.psi~1, locations=train, nmax=nmx, set=list(idp=idp))
  p <- predict(m, newdata=test, debug.level=0)$var1.pred
  RMSE(test$mean.psi, p)
}

idwrasters_occupancy_first<-list()

sppnames<-rownames(phwhall.1316.ordered$y)
saveRDS(sppnames,"sppnames_first.RData")

for(i in 1:length(sppnames)){
  plot.dat.ms <- data.frame(x = phwhall.1316.ordered$coords[,1], 
                            y = phwhall.1316.ordered$coords[,2], 
                            mean.psi = apply(out.alltax.allsigcor.impdet.ms.1316$psi.samples[,i,], 2, mean))
  
  dsp <- SpatialPoints(plot.dat.ms[,1:2])
  dsp <- SpatialPointsDataFrame(dsp, plot.dat.ms)
  crs(dsp)<-CRS("+init=epsg:3734")
  j <- sample(nrow(dsp), 0.2 * nrow(dsp))
  tst <- dsp[j,]
  trn <- dsp[-j,]
opt <- optim(c(8, .5), f1, test=tst, train=trn)
  # define groups for mapping
  cuts <- c(0,0.25,0.75,1.0)
  # set up a palette of interpolated colors
  blues <- colorRampPalette(c('yellow', 'orange', 'blue', 'dark blue'))
  spplot(dsp, 'mean.psi', cuts=cuts, col.regions=blues(5), pch=20, cex=2)
  v <- voronoi(dsp)
   
  ch <- chull(plot.dat.ms[,1:2])
  coords <- plot.dat.ms[c(ch, ch[1]), 1:2]  # closed polygon
  
  plot(plot.dat.ms[,1:2], pch=19)
  lines(coords, col="red")
  
  r <- raster(respoly, res=500)
  
  fitmax <- gstat::gstat(formula = mean.psi ~ 1, data = dsp, nmax = opt$par[1], set = list(idp = opt$par[2]))
  maxint <- raster::interpolate(r, model=fitmax)
  crs(maxint)<-"+proj=lcc +lat_0=39.6666666666667 +lon_0=-82.5 +lat_1=41.7 +lat_2=40.4333333333333 +x_0=600000 +y_0=0 +datum=NAD83 +units=us-ft +no_defs"
  plot(maxint, col=rev(heat.colors(255)))
  plot(vr, add=TRUE)
  r2 <- crop(maxint, extent(respoly))
  r3 <- mask(r2, respoly)
  plot(r3)
  idwrasters_occupancy_first[[i]]<-r3
  writeRaster(r3,filename = paste0("G:/NaturalResources/Byer/DataAnalysis/PHWH/rasterpredictions/02082023_predictions/first/",sppnames[i],"_first_meanpsi_02082023.asc"),overwrite=TRUE)
  
}

saveRDS(idwrasters_occupancy_first,"idwrasters_occupancy_first.RData")

options(max.print=1000000)
sink("out.alltax.allsigcor.impdet.ms.1316.txt")
print(summary(out.alltax.allsigcor.impdet.ms.1316,level="species",quantiles=c(0.05,0.5,0.95)))
sink() 

```


# occupancy model for 1821 alone

```{r, eval = FALSE, warning = FALSE, message= FALSE}

phwhall_occobject_1821_2<-phwhall_occobject_1821

phwhall_occobject_1821_2$y<-phwhall_occobject_1821$y[(apply(phwhall_occobject_1821$y, 1, mean, na.rm = TRUE) > 0),,]
dim(phwhall_occobject_1821_2$y)


sppnames_1821_2<-rownames(phwhall_occobject_1821_2$y)

sort(apply(phwhall_occobject_1821_2$y, 1, mean, na.rm = TRUE))

# louvain's cluster identified 3 clusters - one of (perhaps) more rocky stream communities, one with most fish that are indicators of # stream health, and another of more spring-associated communities. 

#Place a common species first. The first species has all of its factor loadings set to fixed values, and so it can have a large #influence on the resulting interpretation on the factor loadings and latent factors. We have also found that having a very rare #species first can result in very slow mixing and increased sensitivity to initial values of the latent factor loadings matrix.
#For the remaining q1 factors, place species that you believe will show different occurrence patterns than the first species, and #the other species placed before it. Place these remaining q1 species in order of decreasing differences from the initial factor.

#For example, if I had q=3, for the second species in the array, I would place a species that I a priori think is most different from #the first species. For the third species in the array, I would place a species that I think will show different occurrence patterns #than both the first and second species, but its patterns may not be as noticeably different compared to the first and second species.


# Reorder species. 
sp.1821.ordered <- c("midges","streambiotic_suff_mole",sppnames_1821_2[!(sppnames_1821_2 %in% c("midges","streambiotic_suff_mole"))])

# Create new detection-nondetection data matrix in the new order
y.1821.new <- phwhall_occobject_1821_2$y[sp.1821.ordered, , ]
# Create a new data array
phwhall.1821.ordered <- phwhall_occobject_1821_2
# Change the data to the new ordered data
phwhall.1821.ordered$y <- y.1821.new
str(phwhall.1821.ordered)

phwhall_occobject_allsigcor.occ.formula <- ~scale(pool_depth_.cm.) + scale(per_sand) + scale(per_cobble) + scale(substrate_metric_total) +
  scale(pool_metric_total)
phwhall_occobject_dayyear.det.formula <- ~scale(dayyear)

# Number of species
N <- dim(phwhall.1821.ordered$y)[1]
# Distances between sites
dist.phwh <- dist(phwhall.1821.ordered$coords)
# Exponential covariance model
cov.model <- "exponential"
ms.inits <- list(alpha.comm = 0, 
                 beta.comm = 0, 
                 beta = 0, 
                 alpha = 0,
                 tau.sq.beta = 1, 
                 tau.sq.alpha = 1, 
                 z = apply(phwhall.1821.ordered$y, c(1, 2), max, na.rm = TRUE))
ms.priors <- list(beta.comm.normal = list(mean = 0, var = 2.72),
                  alpha.comm.normal = list(mean = 0, var = 2.72), 
                  tau.sq.beta.ig = list(a = 0.1, b = 0.1), 
                  tau.sq.alpha.ig = list(a = 0.1, b = 0.1))


out.alltax.allsigcor.impdet.ms.1821 <- msPGOcc(occ.formula = phwhall_occobject_allsigcor.occ.formula, 
                                    det.formula = phwhall_occobject_dayyear.det.formula, 
                                    data = phwhall.1821.ordered, 
                                    inits = ms.inits, 
                                    n.samples = 30000, 
                                    priors = ms.priors, 
                                    n.omp.threads = 4, 
                                    verbose = TRUE, 
                                    n.report = 10000, 
                                    n.burn = 5000,
                                    n.thin = 50, 
                                    n.chains = 3)
summary(out.alltax.allsigcor.impdet.ms.1821, level = 'both')
ppc.alltax.allsigcor.impdet.ms.1821.out <- ppcOcc(out.alltax.allsigcor.impdet.ms.1821, 'chi-squared', group = 1)
summary(ppc.alltax.allsigcor.impdet.ms.1821.out, level = 'both')

waicOcc(out.alltax.allsigcor.impdet.ms.1821)

write_rds(out.alltax.allsigcor.impdet.ms.1821,"alltaxmodel_phwh_1821.rds")

```

## occupancy predictions

```{r, eval = FALSE, warning = FALSE, message= FALSE}
respoly<-readOGR("G:/NaturalResources/Byer/DataAnalysis/bbs_cmp_analyses/bbsapp/CMPboundaries.shp")

RMSE <- function(observed, predicted) {
  sqrt(mean((predicted - observed)^2, na.rm=TRUE))
}
f1 <- function(x, test, train) {
  nmx <- x[1]
  idp <- x[2]
  if (nmx < 1) return(Inf)
  if (idp < .001) return(Inf)
  m <- gstat(formula=mean.psi~1, locations=train, nmax=nmx, set=list(idp=idp))
  p <- predict(m, newdata=test, debug.level=0)$var1.pred
  RMSE(test$mean.psi, p)
}

idwrasters_occupancy_second<-list()

sppnames<-rownames(phwhall.1821.ordered$y)
saveRDS(sppnames,"sppnames_second.RData")

for(i in 1:length(sppnames)){
  plot.dat.ms <- data.frame(x = phwhall.1821.ordered$coords[,1], 
                            y = phwhall.1821.ordered$coords[,2], 
                            mean.psi = apply(out.alltax.allsigcor.impdet.ms.1821$psi.samples[,i,], 2, mean))
  
  dsp <- SpatialPoints(plot.dat.ms[,1:2])
  dsp <- SpatialPointsDataFrame(dsp, plot.dat.ms)
  crs(dsp)<-CRS("+init=epsg:3734")
  j <- sample(nrow(dsp), 0.2 * nrow(dsp))
  tst <- dsp[j,]
  trn <- dsp[-j,]
opt <- optim(c(8, .5), f1, test=tst, train=trn)
  # define groups for mapping
  cuts <- c(0,0.25,0.75,1.0)
  # set up a palette of interpolated colors
  blues <- colorRampPalette(c('yellow', 'orange', 'blue', 'dark blue'))
  spplot(dsp, 'mean.psi', cuts=cuts, col.regions=blues(5), pch=20, cex=2)
  v <- voronoi(dsp)
   
  ch <- chull(plot.dat.ms[,1:2])
  coords <- plot.dat.ms[c(ch, ch[1]), 1:2]  # closed polygon
  
  plot(plot.dat.ms[,1:2], pch=19)
  lines(coords, col="red")
  
  r <- raster(respoly, res=500)
  
  fitmax <- gstat::gstat(formula = mean.psi ~ 1, data = dsp, nmax = opt$par[1], set = list(idp = opt$par[2]))
  maxint <- raster::interpolate(r, model=fitmax)
  crs(maxint)<-"+proj=lcc +lat_0=39.6666666666667 +lon_0=-82.5 +lat_1=41.7 +lat_2=40.4333333333333 +x_0=600000 +y_0=0 +datum=NAD83 +units=us-ft +no_defs"
  plot(maxint, col=rev(heat.colors(255)))
  plot(vr, add=TRUE)
  r2 <- crop(maxint, extent(respoly))
  r3 <- mask(r2, respoly)
  plot(r3)
  idwrasters_occupancy_second[[i]]<-r3
  writeRaster(r3,filename = paste0("G:/NaturalResources/Byer/DataAnalysis/PHWH/rasterpredictions/02082023_predictions/second/",sppnames[i],"_second_meanpsi_02082023.asc"),overwrite=TRUE)
  
}

saveRDS(idwrasters_occupancy_second,"idwrasters_occupancy_second.RData")

options(max.print=1000000)
sink("out.alltax.allsigcor.impdet.ms.1821.txt")
print(summary(out.alltax.allsigcor.impdet.ms.1821,level="species",quantiles=c(0.05,0.5,0.95)))
sink() 


```